---
title: "HF Trends, 1988 to 2023"
format: 
  html:
    toc: true
    toc-expand: true
    toc-location: body
date: 12-15-2024
theme: lightly
author:
  - name: "Ahmed Sayed (asu.ahmed.sayed@gmail.com)"
    affiliations: "Houston Methodist DeBakey Heart & Vascular Center"
embed-resources: true
toc-title: Contents
page-layout: full
code-fold: true
editor: visual
editor_options: 
  chunk_output_type: console
---

# Setup

## Libraries

```{r}
#Load libs
packs <- c("data.table", "dplyr", "foreach",
           "stringr", "haven", "PooledCohort",
           "ggplot2", "fastsurvey", "scales",
           "flextable", "ggsci", "ggthemes",
           "scales", "mgcv", "Hmisc", "msm", "svrep", "parallel"
           )
lapply(packs, require, character.only = TRUE)
options(survey.lonely.psu = "adjust")
source("multinomial_weighted_mod.R")
```

## Self-made functions

### Analytic

```{r}
#Quantiles
qs <- function(variable) quantile(variable, seq(0, 1, 0.05), na.rm = T)



#Function to rename columns. Please note that this is a more general function not intended solely for this analysis and therefore contains other variables irrelevant to it.
nhanes_rename <- function(x) { x %>% rename(any_of(c(
  #Survey cycle and demographics
  "survey_cycle" = "SDDSRVYR",
  "age" = "RIDAGEYR",
  "sex" = "RIAGENDR",
  "race" = "RIDRETH1",
  #Medical conditions
  "htn_hist" = "BPQ020",
  "dyslipid_hist" = "BPQ080",
  "dm" = "DIQ010",
  "pre_dm" = "DIQ160",
  "chd" = "MCQ160C",
  "angina" = "MCQ160D",
  "mi" = "MCQ160E",
  "stroke" = "MCQ160F",
  "smoke_100" = "SMQ020",
  "smoke_Now" = "SMQ040",
  "cancer" = "MCQ220",
  "dialysis" = "KIQ025",
  "failing_kidney" = "KIQ022",
  "hf" = "MCQ160B",
  "hf_age" = "MCD180B",
  "hf_age" = "MCQ180B",
  "mi_age" = "MCD180E",
  "mi_age" = "MCQ180E",
  #Medical recommendations
  "salt_adv_take" = "MCQ371C",
  "chol_meds_adv" = "BPQ090D",
  "chol_meds_take" = "BPQ100D",
  "htn_meds_adv" = "BPQ040A",
  "htn_meds_take" = "BPQ050A",
  "aspirin_adv" = "RXQ510",
  "aspirin_adv_take" = "RXQ515",
  "aspirin_indep" = "RXQ520",
  #Exam data
  "sbp_1" = "BPXOSY1",
  "sbp_2" = "BPXOSY2",
  "sbp_3" = "BPXOSY3",
  "sbp_1" = "BPXSY1",
  "sbp_2" = "BPXSY2",
  "sbp_3" = "BPXSY3",
  "dbp_1" = "BPXODI1",
  "dbp_2" = "BPXODI2",
  "dbp_3" = "BPXODI3",
  "dbp_1" = "BPXDI1",
  "dbp_2" = "BPXDI2",
  "dbp_3" = "BPXDI3",
  "bmi" = "BMXBMI",
  #Lab values
  "fasting_glucose" = "LBXGLU",
  "hba1c" = "LBXGH",
  "apo_b" = "LBXAPB",
  "ldl" = "LBDLDL",
  "ldl_mh" = "LBDLDLM",
  "ldl_nih" = "LBDLDLN",
  "tg" = "LBXTR",
  "hdl" = "LBDHDD",
  "tc" = "LBXTC",
  "creatinine" = "LBXSCR",
  "creatinine" = "LBDSCR",
  "urine_creatinine" = "URXUCR",
  "urine_albumin" = "URXUMS",
  "urine_albumin" = "URXUMASI",
  "platelets" = "LBXPLTSI"
)))
}


#Import an nhanes file given an NHANES folder and file
import_nhanes <- function(nhanes_pair) {
          
          #State directory of files
          files <- dir(paste0(nhanes_dir, nhanes_pair[1]))
          
          
          #Import files in a given subfolder (lapply is Not needed here, but it is made to be more general in case you need more than 1 file a year)
          lapply(files,
                 function(file) {
                    
                    #Read the file
                    data <- read_xpt(file = paste0(nhanes_dir, 
                                                   nhanes_pair[1], "/", 
                                                   file)
                                     )
                    #Rename columns
                    data <- nhanes_rename(data)
                    
                    #If meds, start command to join medications together
                    if(nhanes_pair[2] == "meds") {
                      
                      #Join medications together
                      if("RXDDRUG" %in% colnames(data)) {
                        data <- data %>%
                          group_by(SEQN) %>%
                          summarise(drug_names = paste(RXDDRUG, collapse = "; ")
                                    )
                        
                      } else if("RXD240B" %in% colnames(data)) {
                        
                        data <- data %>%
                          group_by(SEQN) %>%
                          summarise(drug_names = paste(RXD240B, collapse = "; "),
                                    )
                      }
                    }
                    
                    #Return the object
                    data
                   
                 }
                    
          ) %>%
            #Use rbindlist in case you are importing several years at once
            rbindlist(fill = T) -> data
          
          
}

#Convert variables coded as 1279 in NHANES to Yes/No
convert_1279 <- function(variable, Yes_ind = "Yes", No_ind = "No") {
  fcase(variable == 1, Yes_ind,
        variable == 2, No_ind) %>%
    factor
}

#Convert variables coded as 123479 in NHANES to Yes/No (this is for aspirin use)
convert_123479 <- function(variable, Yes_ind = "Yes", No_ind = "No") {
  fcase(variable %in% c(1, 3), Yes_ind,
        variable %in% c(2, 4), No_ind) %>%
    factor
}


#Convert 77777/99999
convert_5_79 <- function(variable) {
  fifelse(variable %in% c(77777, 99999), NA, variable)
}

convert_3_89 <- function(variable) {
  fifelse(variable %in% c(888, 999), NA, variable)
}

convert_4_89 <- function(variable) {
  fifelse(variable %in% c(8888, 9999), NA, variable)
}


#Convert race
convert_race_complete <-  function(variable) {
  fcase(variable %in% c(1, 2), "Hispanic",
        variable == 3, "Non-Hispanic White",
        variable == 4, "Non-Hispanic Black",
        variable == 6, "Non-Hispanic Asian",
        variable == 7, "Other (including multiracial)"
        )
}

#Convert race (excluding NH Asian; for backward compatability prior to ~2012)
convert_race_incomplete <-  function(variable) {
  fcase(variable %in% c(1, 2), "Hispanic",
        variable == 3, "Non-Hispanic White",
        variable == 4, "Non-Hispanic Black",
        variable == 5, "Other"
        )
}

#Convert NHANES III
convert_race_nhanes_iii <-  function(variable) {
  fcase(variable == 1, "Non-Hispanic White",
        variable == 2, "Non-Hispanic Black"
        )
}

#Function to indicate if drug was taken
find_drug <- function(drugs_column, drug_string, Yes_ind = "Yes", No_ind = "No") {
  fifelse(str_detect(drugs_column, drug_string), Yes_ind, No_ind)
}

#Function to indicate if indication is present
find_ind <- function(inds_column, ind_string, Yes_ind = "Yes", No_ind = "No") {
  fifelse(str_detect(inds_column, ind_string), Yes_ind, No_ind)
}


#Create a function to get survey means and CIs
svymean_df <- function(design, variables) {
  
  #Get chosen variables in the form of "or"
  variables[2] %>%
    #Convert to character
    paste0 %>%
    #Replace + with or
    str_replace_all("\\+", "\\|") %>%
    #Squish together
    str_replace_all(" \\| ", "\\|") -> chosen_vars
  
  #Get the part of svymean's name that corresponds to the variable name
  svymean(design = design, variables, na.rm = T) %>%
    names %>%
    str_extract(chosen_vars) -> chosen_vars
  
  
  #Produce data.table containing the information on totals
  data.table(
    #Variable
    var = chosen_vars,
    #Variable level
    #Total
    pe = svymean(design = design, variables, na.rm = T) %>%
      data.frame %>%
      {.$mean},
    #LCI
    lci = svymean(design = design, variables, na.rm = T) %>% confint %>% data.frame %>% {.$X2.5..},
    #UCI
    uci = svymean(design = design, variables, na.rm = T) %>% confint %>% data.frame %>% {.$X97.5..}
  ) -> mean_df
  
  #Return
  mean_df
  
}


svytotal_df_by <- function(var, by_var, 
                           x = full_data,
                           survey_object = sdata) {
  x[, get(by_var) %>% unique] %>% na.omit -> by_var_levels
  
  #Formula for the by variable
  var_formula <- paste0("~ ", var) %>% as.formula
  
  #Lapply (loop over levels using svytotal_df)
  lapply(by_var_levels,
         function(by_var_level)
           svytotal_df(design = subset(survey_object,
                                       subset = get(by_var) == by_var_level),
                       var_formula) %>%
           mutate(by_var = by_var_level)
         ) %>%
    rbindlist
  
  }


#Create a function to neatly get survey totals/proportions and confidence intervals
svytotal_df <- function(design, variables) {
  
  #Get chosen variables in the form of "or"
  variables[2] %>%
    #Convert to character
    paste0 %>%
    #Replace + with or
    str_replace_all("\\+", "\\|") %>%
    #Squish together
    str_replace_all(" \\| ", "\\|") %>%
    str_remove_all("I\\(|\\(|\\)") -> chosen_vars
  
  svytotal(design = design, variables, na.rm = T) -> svytotal_object
  #Get the part of svytotal's name that corresponds to the variable name
  svytotal_object %>%
    names %>%
    str_extract(chosen_vars) -> chosen_vars
  
  #Get the part of svytotal's name that corresponds to the variable value
  svytotal_object %>%
    names %>%
    str_remove_all(chosen_vars) %>%
    str_remove_all("I\\(|\\(|\\)") -> chosen_vars_levels
  
  #Produce data.table containing the information on totals
  data.table(
    #Variable
    var = chosen_vars,
    #Variable level
    value = chosen_vars_levels,
    #Total
    pe = svytotal_object %>%
      data.frame %>%
      {.$total},
    #LCI
    lci = svytotal_object %>% confint %>% data.frame %>% {.$X2.5..},
    #UCI
    uci = svytotal_object %>% confint %>% data.frame %>% {.$X97.5..}
  ) -> total_df
  
  lapply(chosen_vars_levels,
         function(level) {
           form <- bquote(~ .(variables[[2]]) == .(level))
           #Get proportion
           svyciprop(design = design, 
          formula = form, na.rm = T
          ) %>% {data.table(
               prop_pe = .[1],
               prop_lci = attr(., "ci") %>% .["2.5%"],
               prop_uci = attr(., "ci") %>% .["97.5%"]
             )}
         }
  ) %>%
    rbindlist -> props_df
  
  #Bind
  df <- cbind(total_df, props_df)
  
  #Return
  df
  
}


```

####Props/RRs/Softmax

```{r}


#Softmax function
softmax <- function(logits) {
  exp(c(0, logits))/sum(exp(c(0, logits)))
}


#Function to get multinomial proprotions alongside corresponding confidence intervals
#As input, the function requires a p by k coefficient matrix (k = p = N of coefficients N of outcomes), the variance covariance matrix,
#and a design matrix
#An optional newdata argument can be used if something other than that defined in the main function is needed
get_props_ci <- function(coef_matrix, vcov_matrix, design_matrix, predictors_df = newdata) {
  
  
  #N of outcomes
  k <- ncol(coef_matrix)
  #Multiply by the design matrix to get a matrix of logit predictions
  logit_preds <- design_matrix %*% coef_matrix
  
  #Get CIs for each k outcome per year and place them in a list
  lapply(1:nrow(logit_preds), 
         function(i) {
           
           #Get point estimates
           prop_pe <- softmax(logit_preds[i,])
           
           #Obtain the gradient of the logit
           grad_logit <- kronecker(diag(k), t(design_matrix[i, ]))
           
           #Variance of the logit
           var_logit <- grad_logit %*% vcov_matrix %*% t(grad_logit)
          
           #Symbols to loop over (reference category and all others)
           loop_symbols <- c(0, paste0("x", 1:k))
           denom <- loop_symbols %>% paste0("exp(", ., ")", collapse = "+")
           nums <- loop_symbols %>% paste0("exp(", ., ")")
           
           #Loop over possible k outcomes
           lapply(1:(k + 1),
                  function(j) {
           
                    ##Use delta method to obtain the SE of the proportion on the log scale
                    deltamethod(g = parse(text = paste0("log( ((", nums[j], ")/(", denom, ")) /", 
                                                        "(1 - ((", nums[j], ")/(", denom, ")))", 
                                                        ")" )),
                                mean = logit_preds[i,],
                                cov = var_logit
                    ) -> delta_prop_se
           
           #Get the CIs on the log scale
           plogis(qlogis(prop_pe[j]) + (c(-1.96, 0, 1.96)*delta_prop_se))
           
                  }
           )  -> cis
           
           #Loop over sum of k outcomes (cumulative proportions)
           num <- loop_symbols[-1] %>% paste0("exp(", ., ")") %>% paste0(collapse = "+")
           
           ##Use delta method to obtain the SE of the proportion on the log scale
           deltamethod(g = parse(text = paste0(text = paste0("log( ((", nums, ")/(", denom, ")) /", 
                                                             "(1 - ((", nums, ")/(", denom, ")))", 
                                                             ")" ))),
                       mean = logit_preds[i,],
                       cov = var_logit
           ) -> delta_prop_se
           
           #Get the CIs on the log scale
           plogis(qlogis(sum(prop_pe[-1])) + (c(-1.96, 0, 1.96)*delta_prop_se)) -> cis_cum
           
           #Place CIs into the k+2 element of the list
           cis[[k+2]] <- cis_cum
           
           #Return CIs
           cis
         }
  ) -> pe_ci_list
  

  #Convert list to a data.table with PEs and CIs
  lapply(1:(k + 2), function(k)
    lapply(1:length(pe_ci_list), function(i) {
      
      #Create a data.table
      pe_ci_list[[i]][[k]] %>%
             {data.table(
               lci = 100*.[1],
               pe = 100*.[2],
               uci = 100*.[3]
             )} -> df_year
      
      #Bind with predictor values
      cbind(df_year, data.table(predictors_df)[i,]) -> df_year
      
      #Add pe_ci
      df_year[, pe_ci := paste0(r1(pe), " (", r1(lci), " to ", r1(uci), ")")]
      
    }
      #Bind together yearly predictions for each k data.tables
    ) %>% rbindlist) -> pe_ci_list
  
  #Name the subcomponents of the list
  names(pe_ci_list) <- c("baseline", paste0("category_",(1:k)), "cumulative")
  
  #Return
  pe_ci_list
  
}

#Function to get risk ratios with confidence intervals from a multinomial model
#As input, the function requires a p by k coefficient matrix (k = p = N of coefficients N of outcomes), the variance covariance matrix,
#and a design matrix (2 by p)
get_rr_ci <- function(coef_matrix, vcov_matrix, design_matrix) {
  
  #N of outcomes
  k <- ncol(coef_matrix)
  #Get logits in first and last year being compared
  logit_preds <- design_matrix %*% coef_matrix
  #Get point estimates of risk ratio
  rrs <- softmax(logit_preds[2, ])/softmax(logit_preds[1,])
  rrs[4] <- sum(softmax(logit_preds[2, ])[2:(k+1)])/sum(softmax(logit_preds[1, ])[2:(k+1)])
  
  #Obtain the gradient of the logit
  grad_logit <- lapply(1:k, function(i) kronecker(diag(k), t(design_matrix[i,]))) %>% do.call(rbind, .)
  
  #Variance of the logit
  var_logit <- grad_logit %*% vcov_matrix %*% t(grad_logit)
  
  #Symbols to loop over (reference category and all others)
  loop_symbols1 <- c(0, paste0("x", 1:k))
  loop_symbols2 <- c(0, paste0("x", (k + 1):(2 * k)))
  
  #Numerator/denominator for probability in 1988
  denom1 <- loop_symbols1 %>% paste0("exp(", ., ")", collapse = "+")
  nums1 <- loop_symbols1 %>% paste0("exp(", ., ")")
  
  #Numerator/denominator for probability in 2023
  denom2 <- loop_symbols2 %>% paste0("exp(", ., ")", collapse = "+")
  nums2 <- loop_symbols2 %>% paste0("exp(", ., ")")
  
  
  #Build a vector of means for the delta method:
  #x1 and x2 are 1998; x3 and x4 are 2023
  mean_vector <- c(logit_preds[1,], logit_preds[2,])

  #Loop over categories (& sums of categories) to obtain RRs and their CIs
  lapply(1:(k + 1),
         function(j) {
           
           #P1 and P2
           p1 <- paste0("((", nums1[j], ")/(", denom1, "))" )
           p2 <- paste0("((", nums2[j], ")/(", denom2, "))" )
           rr_exp <- parse(text = paste0("log(", p2, "/", p1, ")" ))
           
           #Delta method
           deltamethod(g = ~ log(
             ( (exp(x4)) /(1 + exp(x3) + exp(x4)))
             /
               ( (exp(x2))/(1 + exp(x1) + exp(x2)))  
           ),
           mean = mean_vector,
           cov = var_logit
           ) -> delta_rr_se
           
           #Get PE and CI
           exp(log(rrs[j]) + c(-1.96, 0, 1.96)*delta_rr_se)
           
         }
         ) -> pe_cis
  
  
        #Loop over sum of k outcomes (cumulative proportions)
  #Cumulative proportions in 1988 and 2023
  num1 <- loop_symbols1[-1] %>% paste0("exp(", ., ")") %>% paste0(collapse = "+")
  num2 <- loop_symbols2[-1] %>% paste0("exp(", ., ")") %>% paste0(collapse = "+")
  
  #P1 and P2
  p1 <- paste0("((", num1, ")/(", denom1, "))" )
  p2 <- paste0("((", num2, ")/(", denom2, "))" )
  rr_exp <- parse(text = paste0("log(", p1, "/", p2, ")" ))
  
        ##Use delta method to obtain the SE of the proportion on the log scale
        deltamethod(g = rr_exp,
                    mean = mean_vector,
                    cov = var_logit
        ) -> delta_rr_se
        
        #Get the CIs on the log scale
        exp(log(rrs[k + 2]) + (c(-1.96, 0, 1.96)*delta_rr_se)) -> pe_cis_cum
        
        #Place CIs into the k+2 element of the list
        pe_cis[[k+2]] <- pe_cis_cum
        
        #Convert each part of the list into a data.table
        lapply(pe_cis, function(x) matrix(x, ncol = 3) %>% 
                 data.table %>%
                 setnames(new = .q(lci, pe, uci)) %>%
                 mutate(pe_ci = paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")"))
               ) -> pe_cis_list
        
        #Name list
        names(pe_cis_list) <- c("baseline", paste0("category_",(1:k)), "cumulative")
        
        #Output
        pe_cis_list

  }


```

### Aesthetic

```{r}

#Create PE (CI) column
pe_ci <- function(pe = pe, lci = lci, uci = uci, rounding = 2) {
  paste0(
    pe  %>% round(rounding) %>% format(nsmall = rounding), " (", 
    lci %>% round(rounding) %>% format(nsmall = rounding), " to ",
    uci  %>% round(rounding) %>% format(nsmall = rounding),
    ")"
  )
}


#Loop over numeric variables, sex, and binary (Yes/No) variables and summarize
sum_var <- function(dataset, variable) {
  #Numeric vars 
  if(variable %in% numeric_variables) {
    dataset[, 
      paste0(median(get(variable), na.rm = T) %>% round(1) %>% format(nsmall = 1),
             " (", 
             iqr(get(variable)), ")"
      )
    ]
  } 
  
  #Sex
  else if(variable %>% str_detect("sex|gender")) {
    paste0(dataset[get(variable) %in% c("female", "Female", "Women", "women"), .N] %>% scales::comma(),
           " (", (100*dataset[get(variable) %in% c("female", "Female", "Women", "women"), .N]/dataset[, .N]) %>% r1, "%)")
  }
  
  #Binary vars
  else if(variable %in% binary_variables) {
    paste0(dataset[get(variable) %in% c("Yes", "Yes"), .N] %>% scales::comma(),
           " (", (100*dataset[get(variable) %in% c("Yes", "Yes"), .N]/dataset[, .N]) %>% r1, "%)")
  }
}


#Get IQRs
iqr <- function(x, rounding = 1) {
  paste0(
    quantile(x, 0.25, na.rm = T) %>% round(rounding) %>% format(nsmall = rounding),
    " to ",
    quantile(x, 0.75, na.rm = T) %>% round(rounding) %>% format(nsmall = rounding)
  )
  
}

#Round to 1 decimal and include zero if needed
r1 <- function(x) {
  x %>% round(1) %>% format(nsmall = 1)
}

#Round to 2 decimals and include zero if needed
r2 <- function(x) {
  x %>% round(2) %>% format(nsmall = 2)
}

#Perform some aesthetic modifications for flextables
flextable_aes <- function(table, table_title) {
  
  #Table font
  table <- flextable::font(table, fontname = "Times New Roman", part = "all")
  
  #Align everything centrally
  table <- flextable::align(table, align = "center", part = "all")
  
  #Table heading
  table <- flextable::set_caption(table, caption = table_title)
  
  #Align footer to the left
  table <- flextable::align(table, align = "left", part = "footer")
  
  
  #Set Table width
  table <- flextable::width(table, width = 1.5, unit = "in")
  
  #Bold
  table <- flextable::bold(table, part = "header")
  
  #Borders
  table <- table %>% border_inner %>% border_outer
  
  #Print
  table
}

```

 # NHANES (1999-2023)

## Import

```{r}

#The lefthand side is the name of the folder in which the respective NHANES data files should be deposited. The lefthand side is the name that will be used for the data.frame in R
folders_dfs <- list(
  c("Demographics", "demo"),
  c("LDL_TG_Measure", "ldl_tg_measure"),
  c("TC_measure", "tc_measure"),
  c("HDL_Measure", "hdl_measure"),
  c("Glycohemoglobin", "hba1c"),
  c("Fasting_Glucose", "fasting_glucose"),
  c("BMI", "bmi"),
  c("BP_Measure", "bp_measure"),
  c("BP_Chol", "bp_chol"),
  c("Diabetes", "dm_df"),
  c("Smoke", "smoke_df"),
  c("Urine albumin and creatinine", "urine_alb_cr"),
  c("Medical Conditions", "med_conds"),
  c("Biochem Profile", "biochem"),
  c("Aspirin", "aspirin"),
  c("Meds", "meds"),
  c("CBC", "cbc"),
  c("Kidney", "kidney")
)

#Get NHANES directory
nhanes_dir <- "C:/Ahmed's Stuff/ResearchStuff/NHANES_Data/"

#Lapply to import NHANES data for a given set of variables
lapply(folders_dfs, function(folder) import_nhanes(folder)) -> data

#Use Reduce to successively merge elements of the data list
Reduce(function(x, y) merge(x, y, all = T, by = "SEQN"), data) -> data

##TABLE 1 FUNCTION
table1_create <- function(dataset, variables, table1_title, 
                          strata_var = NULL, 
                          p_values = FALSE,
                          p_values_col_ind = 5,
                          table_width = 3,
                          numeric_variables, binary_variables
) {
  
  #Store variable names
  var_names <- names(variables)
  
  
  #Get overall CCs using the sum_var function
  lapply(var_names, function(variable) sum_var(dataset, variable)) -> overall_values
  
  #Create strata CCs using the same but with an additional layer of looping for strata levels
  if(!is.null(strata_var)) {
    
    #Strata levels  
    strata_levels <- dataset[, get(strata_var) %>% factor %>% levels]
    
    #Apply over levels of strata of interest
    lapply(1:length(strata_levels),
           function(i)
           {
             stratum <- strata_levels[[i]]
             #CCs of level 1
             lapply(var_names, function(variable) sum_var(dataset = dataset[get(strata_var) == stratum], variable)) %>%
               unlist
           }
    ) -> strata_values
    
    
  }
  
  
  #Create data.table containing the relevant variables
  data.table(
    variable_names = variables %>% unlist,
    variable_values = overall_values %>% unlist
  ) -> table1_df
  
  #Create new columns with the corresponding values for each strata
  if(!is.null(strata_var)) {
    for(i in 1:length(strata_values)) {
      table1_df[, paste0("level_", i) := strata_values[[i]]]
      setnames(table1_df, old = paste0("level_", i), new = strata_levels[i])
    }
  }
  
  #Get p-values if necessary
  if(p_values) {
    #Loop over variables to get test statistics
    lapply(var_names, 
           function(var_name) test_var(dataset, variable = var_name, strata_var)) %>% unlist -> pvals
    table1_df[, p_values := pvals]
  }
  
  #Rename and create a flextable
  table1_df %>%
    rename(any_of(c(
      "Variables" = "variable_names",
      "Overall" = "variable_values",
      "P-values*" = "p_values"
    ))) %>%
    flextable %>% 
    flextable_aes(table1_title) %>%
    bold(j = 1) %>%
    width(width = table_width) -> table_flex
  
  #Add P-value if relevant
  if(p_values) {
    table_flex <- footNote(table_flex, 
             ref_symbols = footNote_symbol(1), 
             j = p_values_col_ind,
             value = as_paragraph("P-values were obtained from Wilcoxon test for continuous variables or from a Chi-square test for categorical variables."))
  }
  
  #Produce table
  table_flex
}
```

## Clean & Prep

### Non-Survey Columns

```{r}

#Convert variables coded as 1, 2, 7, or 9
columns_1279 <- c("pre_dm", "dm", "htn_hist", "smoke_100", "dyslipid_hist",
                  "chd", "mi", "stroke", "angina", 
                  "cancer", "htn_meds_adv", "htn_meds_take",
                  "aspirin_adv", "aspirin_indep", 
                  "failing_kidney", "dialysis", "hf")
data[, (columns_1279) := lapply(.SD, convert_1279, "Yes", "No"), 
     .SDcols = columns_1279]

#If patient has not been adviced to take anti HTN meds by physician, make "take HTN meds on advice of physician" (htn_meds_take) a No
data[, htn_meds_take := fifelse(htn_meds_adv == "No", "No", "Yes")]

#Convert sex
data[, sex := fcase(
    sex == 1, "Men",
    sex == 2, "Women"
) %>% factor]

#Build a string for statins
statin_string <- c("(?i)LOVASTATIN|PRAVASTATIN|FLUVASTATIN|ROSUVASTATIN|PITAVASTATIN|ATORVASTATIN|SIMVASTATIN|CERIVASTATIN")
ccb_string <- c("(?i)AMLODIPINE|VERAPAMIL|DILTIAZEM")
acei_string <- c("(?i)LISINOPRIL|ENALAPRIL|CAPTOPRIL|BENAZEPRIL|FOSINOPRIL")
arb_string <- c("(?i)LOSARTAN|VALSARTAN|OLMESARTAN|CANDESARTAN|IRBESARTAN|TELMISARTAN|AZILSARTAN")
thiazide_string <- c("(?i)HYDROCHLOROTHIAZIDE|CHLORTHALIDONE|INDAPAMIDE")
bb_string <- c("(?i)ATENOLOL|METOPROLOL|BISOPROLOL")
htn_string <- paste0(ccb_string, "|", acei_string, "|", arb_string, "|", thiazide_string, "|", bb_string)

#Indicate drug use
data[, statin := find_drug(drug_names, statin_string, "Yes", "No")]
data[, ccb := find_drug(drug_names, ccb_string, "Yes", "No")]
data[, acei := find_drug(drug_names, acei_string, "Yes", "No")]
data[, arb := find_drug(drug_names, arb_string, "Yes", "No")]
data[, thiazide := find_drug(drug_names, thiazide_string, "Yes", "No")]
data[, bb := find_drug(drug_names, bb_string, "Yes", "No")]
data[, htn_drug := find_drug(drug_names, htn_string, "Yes", "No")]
data[, htn_drug %>% factor %>% summary]
data[, htn_meds := fifelse(htn_meds_take == "Yes" | htn_drug == "Yes", "Yes", 
                                "No") %>% factor]


#Convert race
data[, race := convert_race_incomplete(race)]
data[, black := fifelse(race == "Non-Hispanic Black", "Yes", "No")]
data[, hispanic := fifelse(race == "Hispanic", "Yes", "No")]
data[, asian := fifelse(race == "Asian", "Yes", "No")]


#Convert current smoking habits
data[, smoke_Now := fcase(
  smoke_Now == 1, "Yes",
  smoke_Now == 2, "Yes",
  smoke_Now == 3, "No",
  smoke_100 == "No", "No"
)]

#Create obesity
data[, obesity := fifelse(bmi >= 30, "Yes", "No") %>%
       factor]
data[, severe_obesity := fifelse(bmi >= 40, "Yes", "No") %>%
       factor]

#Calcualte avg systolic and diastolic BP
data[, sbp := (sbp_1 + sbp_2 + sbp_3)/3]
data[, dbp := (dbp_1 + dbp_2 + dbp_3)/3]

#Define DM
# data[, dm := fcase(
#   hba1c >= 6.5, "Yes",
#   dm == "Yes", "Yes",
#   dm == "No", "No"
#   )]

#DM
data[, dm := fifelse(
  hba1c >= 6.5, "Yes", "No"
  )]
#PreDM
data[, pre_dm := fifelse(
  hba1c >= 5.7 & hba1c < 6.5, "Yes", "No"
  )]
#PreDM or DM
data[, pre_dm_or_dm := fifelse(
  hba1c >= 5.7, "Yes", "No"
  )]


#Calculate sex-specific constants for eGFR estimation
data[, alpha := fifelse(sex == "Men", -0.302, -0.241)]
data[, k := fifelse(sex == "Men", 0.9, 0.7)]
data[, constant := fifelse(sex == "Men", 1, 1.012)]

#Realibrate creatinine in 1999-2000 (See: https://pubmed.ncbi.nlm.nih.gov/18037092/)
data[survey_cycle == 1, creatinine := 0.147 + (creatinine*1.013)]

#Realibrate creatinine in 2005-2006 (See:https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2005/DataFiles/BIOPRO_D.htm)
data[survey_cycle == 4, creatinine := -0.016 + (creatinine*0.978)]

#Calculate eGFR
data[, egfr :=
    142 *
    (pmin(creatinine/k, 1)^alpha) *
    (pmax(creatinine/k, 1)^-1.2) *
    (0.9938^age) *
    constant
    ]


#Correct urine albumin units from 1999 to 2003

#Create albumin/creatinine ratio
data[urine_albumin == 888888, urine_albumin := NA]
data[urine_creatinine == 88888, urine_creatinine := NA]
#Convert to mg/g (1000 for g to mg for urine creatinine and 10 for dL to L for urine creatinine as wel. 1000/10 = 100)
data[, uacr := 100*urine_albumin/urine_creatinine]

```

### Survey-related Columns

```{r}
#Adjust examination survey weights to reflect proportion of data from each cycle
#Exam
data[survey_cycle != "2017 to 2020", exam_weight := WTMEC2YR] 
data[survey_cycle == 66, exam_weight := WTMECPRP] 
data[is.na(exam_weight), exam_weight := 0]

#Interview
data[survey_cycle != "2017 to 2020", int_weight := WTINT2YR] 
data[survey_cycle == 66, int_weight := WTINTPRP] 
data[is.na(int_weight), int_weight := 0]

#Fasting
data[survey_cycle != "2017 to 2020", fast_weight := WTSAF2YR.y] 
data[survey_cycle == 66, fast_weight := WTSAFPRP.x] 
data[is.na(fast_weight), fast_weight := 0]


#Do the same for questionnaire survey weights

#Make last survey follow the previous ones in terms of numeric coding.
data[, survey_cycle := fifelse(survey_cycle == 66, 10, survey_cycle) %>% factor]

#Modify survey_cycle variable
levels(data$survey_cycle) <- c("1999 to 2000", "2001 to 2002", "2003 to 2004", "2005 to 2006", "2007 to 2008", "2009 to 2010", "2011 to 2012", "2013 to 2014", "2015 to 2016", "2017 to 2020", "2021 to 2023")

```

# NHANES III (1988-1994)

## Import

```{r}

#NHANES III has a different file structure to other NHANES files. Therefore, it will be processed using separate blocks of code
#Please note that this NHANES III dataset has been preprocessed. It was originally composed of 4 separate files which were merged together using the SEQN column. The code needed to merge these 4 separate files is provided as a separate document on this page.
library(qs)
qread(paste0(nhanes_dir, "NHANES III/", "NHANES III List.RData")) -> nhanes_iii_data
names(nhanes_iii_data) <- c("adult", "youth", "exam", "lab")

##Remove all column names from other files that overlap with column names in the exam data.frame (except SEQN, which will be used to join them)
colnames(nhanes_iii_data$exam) -> exam_colnames
exam_colnames <- exam_colnames[exam_colnames != "SEQN"]
nhanes_iii_data$adult[, (exam_colnames) := NULL]
nhanes_iii_data$youth[, (exam_colnames) := NULL]
nhanes_iii_data$lab[, (exam_colnames) := NULL]

#Use Reduce to successively merge elements of the data list

# merge(nhanes_iii_data$exam,
#       merge(nhanes_iii_data$lab,
#      rbind(nhanes_iii_data$adult, nhanes_iii_data$youth, fill = T),
#      by = "SEQN",
#      all = T),
#      by = "SEQN",
#      all = T
#      ) -> nhanes_iii_data

Reduce(function(x, y) merge(x, y, all = T, by = "SEQN"), 
    nhanes_iii_data) -> nhanes_iii_data
```

## Clean & Prep

###Prescription data

```{r}

haven::read_xpt(paste0(nhanes_dir, "NHANES III/", "pupremed.xpt")) -> nhanes_iii_prescription_data


nhanes_iii_prescription_data <- data.table(nhanes_iii_prescription_data)
nhanes_iii_prescription_data <- nhanes_iii_prescription_data[, .(HQRXCODE = paste0(HQRXCODE, collapse = ";")), by = "SEQN"]


#Build a string for statins
statin_nhanes_iii_string <- c("1664|3064|3088")

#CCBs, ACEIs, Thiazides
ccb_nhanes_iii_string <- c("3072|2425|1069")
acei_nhanes_iii_string <- c("1647|1458|1138|1139|0642|0643|3053|3052")
#arb_nhanes_iii_string None exists
thiazide_nhanes_iii_string <- c("0180|0643|1139|1451|1452|1456|1457|1458|1459|1460|1461|1462|1463|1464|1465")
bb_nhanes_iii_string <- c("0333|0334|1460|1801")
htn_nhanes_iii_string <- paste0(ccb_nhanes_iii_string, "|", acei_nhanes_iii_string, "|", thiazide_nhanes_iii_string, "|", bb_nhanes_iii_string)

nhanes_iii_prescription_data[, statin := find_drug(HQRXCODE, statin_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[statin == "Yes", .N]
nhanes_iii_prescription_data[, ccb := find_drug(HQRXCODE, ccb_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[, acei := find_drug(HQRXCODE, acei_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[, arb := "No"]
nhanes_iii_prescription_data[, thiazide := find_drug(HQRXCODE, thiazide_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[, bb := find_drug(HQRXCODE, bb_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[, htn_drug := find_drug(HQRXCODE, htn_nhanes_iii_string, "Yes", "No")]
nhanes_iii_prescription_data[htn_drug == "Yes", .N]

#Merge
nhanes_iii_data <- merge(nhanes_iii_data, nhanes_iii_prescription_data, 
                         by = "SEQN", all.x = T, all.y = F)


```

### Non-Survey Columns

```{r}
nhanes_iii_rename <- function(x) { x %>% rename(any_of(c(
  #Survey cycle and demographics
  "survey_cycle" = "SDDSRVYR",
  "age" = "HSAGEIR",
  "age_unit" = "RIDAGEYR",
  "sex" = "HSSEX",
  "sex" = "HSSEX",
  "race" = "DMARACER",
  "ethn" = "DMAETHNR",
  "int_weight1" = "WTPFQX1",
  "int_weight2" = "WTPFQX2",
  "mec_exam_weight1" = "WTPFEX1",
  "mec_exam_weight2" = "WTPFEX2",
  "mh_exam_weight1" = "WTPFHX1",
  "mh_exam_weight2" = "WTPFHX2",
  "morning_weight1" = "WTPFHSD1",
  "morning_weight2" = "WTPFHSD2",
  "int_weight1" = "WTPFEX1",
  "int_weight2" = "WTPFEX2",
  "fast_hrs" = "PHPFAST",
  #Medical conditions
  "htn_hist" = "HAE2",
  "dyslipid_hist" = "HAE7",
  "dm" = "HAD1",
  "pre_dm" = "ABSENT",
  "chd" = "ABSENT",
  "angina" = "ABSENT",
  "mi" = "HAF10",
  "mi_age" = "HAF12R",
  "stroke" = "HAC1D",
  "smoke_100" = "SMQ020",
  "smoke_Now" = "SMQ040",
  "cancer" = "HAC10",
  "dialysis" = "ABSENT",
  "failing_kidney" = "ABSENT",
  "hf" = "HAC1C",
  "hf_age" = "HAC3CR",
  #Medical recommendations
  "salt_adv_take" = "HAE5C",
  "chol_meds_adv" = "BPQ090D",
  "chol_meds_take" = "BPQ100D",
  "htn_meds_adv" = "HAE4A",
  "htn_meds_take" = "HAE5A",
  "aspirin_adv" = "ABSENT",
  "aspirin_adv_take" = "ABSENT",
  "aspirin_indep" = "ABSENT",
  #Exam data
  "sbp_1" = "HAZA8AK1",
  "sbp_2" = "HAZA8BK1",
  "sbp_3" = "HAZA8CK1",
  "sbp_4" = "HAZA8DK1",
  "dbp_1" = "HAZA8AK5",
  "dbp_2" = "HAZA8BK5",
  "dbp_3" = "HAZA8CK5",
  "dbp_4" = "HAZA8DK5",
  "height" = "BMPHT",
  "body_weight" = "BMPWT",
  #Lab values
  "fasting_glucose" = "ABSENT",
  "hba1c" = "GHP",
  "apo_b" = "ABP",
  "ldl" = "LCP",
  "tg" = "TGP",
  "hdl" = "HDP",
  "tc" = "TCP",
  "creatinine" = "CEP",
  "urine_creatinine" = "URP",
  "urine_albumin" = "UBP",
  "platelets" = "PLP"
)))
}

#Remove .x
colnames(nhanes_iii_data) <- colnames(nhanes_iii_data) %>%
  str_remove(".x")

#Keep unique columns only
nhanes_iii_data[, .SD, .SDcols = unique(names(nhanes_iii_data))] -> nhanes_iii_data

#Rename columns
nhanes_iii_rename(nhanes_iii_data) -> nhanes_iii_data

#Convert variables coded as 1, 2, 7, or 9
nhanes_iii_columns_1279 <- columns_1279[columns_1279 %in% names(nhanes_iii_data)]
nhanes_iii_data[, 
                (nhanes_iii_columns_1279) := lapply(.SD, convert_1279, "Yes", "No"), 
     .SDcols = nhanes_iii_columns_1279]

#If patient has not been adviced to take anti HTN meds by physician, make "take HTN meds on advice of physician" (htn_meds_take) a No
nhanes_iii_data[, htn_meds_take := fifelse(htn_meds_adv == "No", "No", "Yes")]
nhanes_iii_data[, htn_meds := fifelse(htn_meds_take == "Yes" | htn_drug == "Yes", "Yes", 
                                "No") %>% factor]


#NA absent columns
nhanes_iii_data[body_weight == 888888, body_weight := NA]
nhanes_iii_data[height == 88888, height := NA]
nhanes_iii_data[creatinine == 8888, creatinine := NA]
nhanes_iii_data[fast_hrs == 88888, fast_hrs := NA]
#NA tgs if non-fasting
nhanes_iii_data[fast_hrs < 9 | is.na(fast_hrs), tg := NA]



#Convert numerically coded NA SBP/DBP
nhanes_iii_data[, names(.SD) := lapply(.SD, convert_3_89),
                .SDcols = c(paste0("sbp_", 1:3))]
nhanes_iii_data[, names(.SD) := lapply(.SD, convert_3_89),
                .SDcols = c(paste0("dbp_", 1:3))]
#Calcualte avg systolic and diastolic BP
nhanes_iii_data[, sbp := (sbp_1 + sbp_2 + sbp_3)/3]
nhanes_iii_data[, dbp := (dbp_1 + dbp_2 + dbp_3)/3]


#Convert sex
nhanes_iii_data[, sex := fcase(
  sex == 1, "Men",
  sex == 2, "Women"
) %>% factor]

#Race and ethnicity
nhanes_iii_data[, race := convert_race_nhanes_iii(race)]
nhanes_iii_data[ethn %in% c(1, 2), race := "Hispanic"]
nhanes_iii_data[is.na(race), race := "Other"]

#Calculate sex-specific constants for eGFR estimation
nhanes_iii_data[, alpha := fifelse(sex == "Men", -0.302, -0.241)]
nhanes_iii_data[, k := fifelse(sex == "Men", 0.9, 0.7)]
nhanes_iii_data[, constant := fifelse(sex == "Men", 1, 1.012)]

#Convert the numerical NAs in Hba1c
nhanes_iii_data[, hba1c := convert_4_89(hba1c)]

#Define DM
# nhanes_iii_data[, dm := fcase(
#   hba1c >= 6.5, "Yes",
#   dm == "Yes", "Yes",
#   dm == "No", "No"
# ) %>% factor]

#DM
nhanes_iii_data[, dm := fifelse(
  hba1c >= 6.5, "Yes", "No"
) %>% factor]
#PreDM
nhanes_iii_data[, pre_dm := fifelse(
  hba1c >= 5.7 & hba1c < 6.5, "Yes", "No"
  ) %>% factor]
#PreDM or DM
nhanes_iii_data[, pre_dm_or_dm := fifelse(
  hba1c >= 5.7, "Yes", "No"
  ) %>% factor]


#Calibrate creatinine (See: https://pubmed.ncbi.nlm.nih.gov/18037092/)
nhanes_iii_data[, creatinine := -0.184 + (0.960*creatinine)]
#Calculate eGFR
nhanes_iii_data[, egfr :=
       142 *
       (pmin(creatinine/k, 1)^alpha) *
       (pmax(creatinine/k, 1)^-1.2) *
       (0.9938^age) *
       constant
]

#Create albumin/creatinine ratio
nhanes_iii_data[urine_albumin == 888888, urine_albumin := NA]
nhanes_iii_data[urine_creatinine == 88888, urine_creatinine := NA]
nhanes_iii_data[, uacr := 100*urine_albumin/urine_creatinine]

#Create BMI
nhanes_iii_data[, bmi :=
                  (body_weight)/
                  ((height/100)^2)
                  ]

#Create obesity
nhanes_iii_data[, obesity := fifelse(bmi >= 30, "Yes", "No") %>%
                  factor]
nhanes_iii_data[, severe_obesity := fifelse(bmi >= 40, "Yes", "No") %>%
                  factor]


```

### Survey-related Columns

```{r}
#Create survey cycle variable
nhanes_iii_data[, survey_cycle := fcase(SDPPHASE == 1, "1988 to 1991",
                                        SDPPHASE == 2, "1991 to 1994") %>% factor]

#Weight
nhanes_iii_data[, exam_weight := fcase(survey_cycle == "1988 to 1991",
                                       mh_exam_weight1,
                                       survey_cycle == "1991 to 1994",
                                       mh_exam_weight2)]

nhanes_iii_data[, int_weight := fcase(survey_cycle == "1988 to 1991",
                                       int_weight1,
                                       survey_cycle == "1991 to 1994",
                                       int_weight2)]
nhanes_iii_data[, fast_weight := fcase(survey_cycle == "1988 to 1991",
                                       morning_weight1,
                                       survey_cycle == "1991 to 1994",
                                       morning_weight2)]
nhanes_iii_data[is.na(fast_weight), fast_weight := 0]

#PSU
nhanes_iii_data[, SDMVPSU := fcase(survey_cycle == "1988 to 1991",
                                       SDPPSU1,
                                       survey_cycle == "1991 to 1994",
                                       SDPPSU2)]

#Strata
nhanes_iii_data[, SDMVSTRA := fcase(survey_cycle == "1988 to 1991",
                                       SDPSTRA1,
                                       survey_cycle == "1991 to 1994",
                                       SDPSTRA2)]


nhanes_iii_data <- nhanes_iii_data[!is.na(exam_weight) | !is.na(int_weight)]
```

# Full NHANES

## Merge

```{r}
#Get columns to keep
colnames(nhanes_iii_data)[colnames(nhanes_iii_data) %in% colnames(data)] -> nhanes_iii_keep
nhanes_iii_data <- nhanes_iii_data[, ..nhanes_iii_keep]

#Merge
rbind(nhanes_iii_data, data, fill = T) -> full_data


```

## Inclusion Criteria

```{r}
#Heart failure, N
hf_n <- full_data[hf == "Yes", .N]
```

## Create/modify some columns

```{r}
full_data[,
          year :=
            fcase(
              survey_cycle == "1988 to 1991", 1990.25,
              survey_cycle == "1991 to 1994", 1993.25,
              survey_cycle == "1999 to 2000", 2000,
              survey_cycle == "2001 to 2002", 2002,
              survey_cycle == "2003 to 2004", 2004,
              survey_cycle == "2005 to 2006", 2006,
              survey_cycle == "2007 to 2008", 2008,
              survey_cycle == "2009 to 2010", 2010,
              survey_cycle == "2011 to 2012", 2012,
              survey_cycle == "2013 to 2014", 2014,
              survey_cycle == "2015 to 2016", 2016,
              survey_cycle == "2017 to 2020", 2018.6,
              survey_cycle == "2021 to 2023", 2022.7
            )
            ]

#Correctly label missing codes for TG/LDL
full_data[, tg := convert_3_89(tg)]
full_data[, ldl := convert_3_89(ldl)]
full_data[, tc := convert_3_89(tc)]
full_data[, tg := convert_4_89(tg)]
full_data[, ldl := convert_4_89(ldl)]

#Cap age at 80 to be consistent throughout
full_data[, age := fifelse(age > 80, 80, age)]

#Column for dyslipidemia
full_data[, dyslipid := fifelse(tg >= 200 | ldl >= 160, "Yes", "No") %>% factor]
full_data[, hyperldl := fifelse(ldl >= 160, "Yes", "No") %>% factor]
full_data[, hypertg := fifelse(tg >= 200, "Yes", "No") %>% factor]
full_data[, hypercholesterol := fifelse(tc >= 200, "Yes", "No") %>% factor]

#Create ordinal sex/race variables (this only serves to estimate a difference-in-smooths GAM later on when we want to test for sex-specific time trends).
full_data[, sex_ord := factor(sex, ordered = T)]
full_data[, race_binary := factor(fifelse(race == "Non-Hispanic White", "Non-Hispanic White", "Non-Hispanic Black, Hispanic, or Other Minority"), ordered = F)]
full_data[, race_binary_ord := factor(fifelse(race == "Non-Hispanic White", "Non-Hispanic White", "Non-Hispanic Black, Hispanic, or Other Minority"), ordered = T)]

#HF as an ordinal variable (same reason as above; to facilitate GAM interaction)
full_data[, hf_ord := factor(hf, ordered = T)]

#Modify columns where DBP is not recorded (missing Korotokoff sound)
full_data[dbp == 0, dbp := NA]

#Create HTN categories
full_data[, htn := fifelse(sbp >= 130 | dbp >= 80, "Yes", "No") %>% factor]
full_data[, htn_stage2 := fifelse(sbp >= 140 | dbp >= 90, "Yes", "No") %>% factor]

#Column for diabetes categorization (for multinomial regression)
full_data[, dm_cats := fcase(
  hba1c >= 6.5, 2,
  hba1c %between% c(5.7, 6.49), 1,
  hba1c < 5.7, 0
)]

full_data[, obesity_cats := fcase(
  bmi >= 40, 2,
  bmi %between% c(30, 39.999), 1,
  bmi < 30, 0
)]

full_data[, htn_cats := fcase(
  sbp >= 140 | dbp >= 90, 2,
  sbp %between% c(130, 139.999) | dbp %between% c(80, 89.999) , 1,
  sbp < 130 & dbp <80, 0
)]


#CKD
full_data[, ckd45 := fifelse(egfr < 30, "Yes", "No") %>% factor]
full_data[, ckd345 := fifelse(egfr < 60, "Yes", "No") %>% factor]
full_data[, ckd345_albumin := fifelse(egfr < 60 | uacr >= 30, "Yes", "No") %>% factor]
full_data[, microalbuminuria := fifelse(uacr %between% c(30, 300), "Yes", "No") %>% factor]
full_data[, macroalbuminuria := fifelse(uacr > 300, "Yes", "No") %>% factor]
full_data[, albuminuria := fifelse(uacr >= 30, "Yes", "No") %>% factor]


#Relevel hypertension factors
full_data[, htn := factor(htn, levels = c("No", "Yes"))]

#Create history of hypertension variable based on present elevation or previous self-reported history
full_data[, htn_hist2 := fifelse(htn_hist == "Yes" | sbp >= 130 | dbp >= 80, "Yes", "No") %>% factor]


#Create history of dyslipidemia variable based on present elevation or previous self-reported history
full_data[, dyslipid_hist2 := fifelse(dyslipid_hist == "Yes" | tc >= 200, "Yes", "No") %>% factor]
full_data[, acei_arb := fifelse(acei == "Yes" | arb == "Yes", "Yes", "No")]
full_data[, htn_meds_count := ((thiazide == "Yes") + (ccb == "Yes") + (bb == "Yes") + (acei_arb == "Yes") )]
full_data[, htn_meds_count_cats := fifelse(htn_meds_count > 2, 2, htn_meds_count)]
full_data[, 
          c("ccb", "bb", "acei_arb", "thiazide", "statin") := lapply(.SD, factor), 
          .SDcols = c("ccb", "bb", "acei_arb", "thiazide", "statin")]


```

## Convert to a survey object

```{r}
#Make survey object out of x

#Exams
sdata <- svydesign(data = full_data,
                  ids = full_data$SDMVPSU,
                strata = full_data$SDMVSTRA,
                weights = full_data$exam_weight,
                nest = TRUE)
sdata <- subset(sdata, age >= 20)

#Questionnaire
sdata_q <- svydesign(data = full_data,
                ids = full_data$SDMVPSU,
                strata = full_data$SDMVSTRA,
                weights = full_data$int_weight,
                nest = TRUE)
sdata_q <- subset(sdata_q, age >= 20)
full_data <- full_data[age >= 20]

#Fasting
sdata_fast <- svydesign(data = full_data,
                ids = full_data$SDMVPSU,
                strata = full_data$SDMVSTRA,
                weights = full_data$fast_weight,
                nest = TRUE)
sdata_fast <- subset(sdata_fast, age >= 20)

full_data <- full_data[age >= 20]
```

# Analyze

## Bootstrap replicates of survey

```{r}
#These two objects will be used throughout the analysis to carry out the bootstrap replicates-based analyses
set.seed(100)
sdata_rep <- as_bootstrap_design(sdata,
                                 type = "Rao-Wu-Yue-Beaumont",
                                 replicates = 500)

sdata_rep_q <- as_bootstrap_design(sdata_q,
                                 type = "Rao-Wu-Yue-Beaumont",
                                 replicates = 500)


sdata_rep_fast <- as_bootstrap_design(sdata_fast,
                                 type = "Rao-Wu-Yue-Beaumont",
                                 replicates = 500)

```

## GAM functions

### \[Binomial outcomes\]

#### Overall

```{r}


# survey_design = sdata_q
# survey_replicates <- sdata_rep_q
# data = full_data
# newdata = newdata_overall
# survey_weights = full_data[hf == "Yes", int_weight]
# int_var <- "overall"
# model_formula <- paste0("pre_dm_or_dm", c(overall_form, sex_form, race_form))[1]
# model_formula <- dyslipid~s(year, k = gam_k)
# 
# rm(survey_design, data, newdata, survey_weights, int_var, model_formula)

#Function to automate GAM estimation for risk factors underlying HF
#int_var is only defined here for the sake of compatability with the subgroup-specific function because they will both be used in the same lapply loop
gam_rep <- function(survey_design,
                    survey_replicates,
                    data,
                    newdata,
                    survey_weights,
                    model_formula,
                    int_var = NULL,
                    return_gam_object = FALSE) {
  
  #Assign survey weights to the data.frame (so that gam finds it)
  data[, survey_weights := survey_weights ]
  #GAM object (to retrieve spline basis matrix)
  gam(
    data = data,
    family = binomial,
    method = "REML",
    formula = model_formula,
    weights = survey_weights) -> gam_object
  
  #Obtain design matrix
  design_matrix <- predict(gam_object, newdata = newdata,
                           type = "lpmatrix")
  
  
  #Get bootstrap estiamte of the variance
  withReplicates(design = survey_replicates,
                 theta = function(w, d) {
                   
                   #Reassign formulas to the current environment
                   f <- model_formula
                   environment(f) <- environment()
                   
                   
                   coef(gam(
                   family = binomial,
                   method = "REML",
                   formula = f,
                   weights = w,
                   data = d)
                 )
                 }
  ) -> reps
  
  
  #Get standard errors and point estimates
  cov_matrix <- design_matrix %*% vcov(reps) %*% t(design_matrix)
  se <- sqrt(diag(cov_matrix))
  pe <- design_matrix %*% coef(reps)
  
  #Store predictions
  data.table(cbind(newdata,
        data.table(
    pe = 100*as.numeric(plogis(pe)),
    lci = 100*as.numeric(plogis(pe - (1.96*se))),
    uci = 100*as.numeric(plogis(pe + (1.96*se)))
  ))) -> preds_df
  
  #Get PE/CI
  preds_df[, pe_ci := paste0(r1(pe), " (", r1(lci), " to ", r1(uci), ")")]
  
  #Apply delta method to obtain SEs of risk ratios
  msm::deltamethod(g = ~ log((exp(x2)/(1+exp(x2))) / (exp(x1)/(1+exp(x1)))),
              mean = pe[c(1, length(pe))],
              cov = cov_matrix[c(1, length(pe)),c(1, length(pe))]
              ) -> rr_se

  #Obtain log-RR
  log((exp(pe[length(pe)])/(1 + exp(pe[length(pe)])))/(exp(pe[1])/(1 + exp(pe[1])))) -> rr_pe
 
  
  #Store RR contrasts
  data.table(pe = exp(rr_pe),
             lci = exp(rr_pe - (1.96*rr_se)),
             uci = exp(rr_pe + (1.96*rr_se))
             ) -> contrasts_df
  #Get PE/CI
  contrasts_df[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]
  
  
  #Get year-wise estimates (no model smoothing)
svytotal_df_by(var = as.character(model_formula[[2]]), 
               by_var = "year",
               x = data,
               survey_object = survey_design) %>%
  filter(value == "Yes") -> year_df
  
#Return gam_object if true
if(!return_gam_object) {gam_object <- NULL}

  #Get output
  list(preds_df = preds_df,
       year_df = year_df,
       contrasts_df = contrasts_df,
       vcov = vcov(reps),
       coef = coef(reps),
       gam_object = gam_object)
  
}

```

#### Interaction

```{r}

# i = 3
# survey_design = sdata %>% subset(hf == "Yes")
# data = full_data[hf == "Yes"]
# newdata = newdata_race
# survey_weights = full_data[hf == "Yes", int_weight]
# int_var <- "race_binary"
# model_formula <- rfs_base_list$model_formulae[3]
# model_formula <- rfs_base_list$model_formulae[[i]] %>% formula
# survey_replicates <- sdata_rep %>% subset(hf == "Yes")

rm(survey_design, data, newdata, survey_weights, int_var, model_formula, survey_replicates)
#Function to automate GAM estimation for risk factors underlying HF
gam_rep_int <- function(survey_design,
                    data,
                    newdata,
                    survey_weights,
                    model_formula,
                    survey_replicates,
                    int_var) {
  
  
  #Assign survey weights to the data.frame (so that gam finds it)
  data[, survey_weights := survey_weights ]
  #GAM object (to retrieve spline basis matrix)
  gam(
    data = data,
    family = binomial,
    method = "REML",
    formula = model_formula,
    weights = survey_weights) -> gam_object
  
  #Obtain design matrix
  design_matrix <- predict(gam_object, newdata = newdata, type = "lpmatrix")
  
  #Get bootstrap estimate of the variance
  withReplicates(design = survey_replicates,
                 theta = function(w, d) {
                   
                   #Reassign formulas to the current environment
                   environment(model_formula) <- environment()
                   
                   
                   coef(gam(
                   family = binomial,
                   method = "REML",
                   formula = model_formula,
                   weights = w,
                   data = d)
                 )
                 }
  ) -> reps
  
  #Get standard errors and point estimates
  cov_matrix <- design_matrix %*% vcov(reps) %*% t(design_matrix)
  se <- sqrt(diag(cov_matrix))
  pe <- design_matrix %*% coef(reps)
  
  #Store predictions
  #Store predictions
  data.table(cbind(newdata,
        data.table(
    pe = 100*as.numeric(plogis(pe)),
    lci = 100*as.numeric(plogis(pe - (1.96*se))),
    uci = 100*as.numeric(plogis(pe + (1.96*se)))
  ))) -> preds_df
  
  #Get PE/CI
  preds_df[, pe_ci := paste0(r1(pe), " (", r1(lci), " to ", r1(uci), ")")]
  
  
  #Get number of unique factors to loop over
  preds_df[, get(int_var) %>% unique] -> unique_factors
  
  #Apply delta method to obtain SEs of risk ratios. 
  lapply(unique_factors,
         function(level) {
           
           #Get the row indices which correspond to the first and last year for each factor
  preds_df[, range(which(get(int_var) == level))] -> level_indices
           
           #Apply the delta method
  msm::deltamethod(g = ~ log((exp(x2)/(1+exp(x2))) / (exp(x1)/(1+exp(x1)))),
                   mean = pe[level_indices],
                   cov = cov_matrix[level_indices, level_indices]
  ) -> rr_se
  
  #Obtain log-RR
  log(preds_df[max(level_indices), pe]/preds_df[min(level_indices), pe]) -> rr_pe
  
  #Store RR contrasts
  data.table(pe = exp(rr_pe),
             lci = exp(rr_pe - (1.96*rr_se)),
             uci = exp(rr_pe + (1.96*rr_se))) -> contrast_df
             
             #Get PE/CI
             contrast_df[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]
             
             #Return
             contrast_df
  }
  ) -> contrasts_df
  
  #Convert to a data.table and store factor levels
  contrasts_df <- rbindlist(contrasts_df)[, int_factor := unique_factors]
  
  #Compare increases
  preds_df[, lapply(unique_factors, 
                    function(level) range(which(get(int_var) == level))) %>%
             unlist %>%
             sort] -> all_levels_indices
  
  #Ratio of ratios
  contrasts_df[, log(max(pe)) - log(min(pe))] -> rrr_pe
  
  #SE for ratio of ratios using the delta method
  msm::deltamethod(g = ~ log((exp(x2)/(1+exp(x2))) / (exp(x1)/(1+exp(x1)))) -
                     log((exp(x4)/(1+exp(x4))) / (exp(x3)/(1+exp(x3)))),
                   mean = pe[all_levels_indices],
                   cov = cov_matrix[
                     all_levels_indices, 
                     all_levels_indices]
  ) -> rrr_se
  
  
  #Identify interaction indices (containing :)
  which(str_detect(names(gam_object$coefficients), ":")) -> selection_vector
  #P-value for interaction
  mgcv:::testStat(p = coef(reps)[selection_vector], V = vcov(reps)[selection_vector, selection_vector],
                X = gam_object$R[, selection_vector],
                rank = gam_object$edf1[selection_vector] %>% sum,
                type = 0,
                res.df = -1)$pval -> p_int
  
  #Create a DF combining them
  rrr_df <- data.table(pe = exp(rrr_pe),
                       lci = exp(rrr_pe - (1.96*rrr_se)),
                       uci = exp(rrr_pe + (1.96*rrr_se)),
                       #P-value for relative difference of 1988 vs 2023 in males vs females (a double contrast in time and sex, at 1988 and 2023 specifically)
                       pval_contrast = 2*(1 - pnorm(abs(rrr_pe), mean = 0, sd = rrr_se)),
                       #P-value for interaction (not for a particular contrast)
                       pval_int = p_int
                       )
  
  
  #Get year-wise estimates (no model smoothing)
  lapply(unique_factors,
         function(level) {
           
           #Subset expression
           subset_survey <- do.call("subset", 
                                      list(x = survey_design, 
                                           subset = parse(text = paste0(int_var, " == \"", level, "\""))
                                           )
                                    )
     
           #year_df
  svytotal_df_by(var = as.character(model_formula[[2]]), 
                 by_var = "year",
                 x = data,
                 survey_object = subset_survey) %>%
  filter(value == "Yes") -> year_df
  
  #Assign name of factor
  year_df[, int_factor := level]
  
  #Return
  year_df
         }
  ) -> year_list
  
  #Convert to a data.table and store factor levels
  year_df <- rbindlist(year_list)
  

  #Get output
  list(preds_df = preds_df,
       year_df = year_df,
       contrasts_df = contrasts_df,
       rrr_df = rrr_df)
  
}
```

### \[Multinomial outcomes\]

#### Overall

```{r}

# survey_design = sdata %>% subset(hf == "Yes" & htn_hist2 == "Yes" & year <= 2020)
# survey_replicates <- sdata_rep %>% subset(hf == "Yes" & htn_hist2 == "Yes" & year <= 2020)
# data = full_data[hf == "Yes" & htn_hist2 == "Yes" & year <= 2020]
# newdata = newdata_overall %>% filter(year <= 2020)
# survey_weights = full_data[hf == "Yes" & htn_hist2 == "Yes" & year <= 2020, int_weight]
# int_var <- "overall"
# model_formula <- rfs_base_list$model_formulae[[1]] %>% lapply(formula)
# constituent_outcomes <- rfs_base_list$constituent_outcomes
# rm(survey_design, survey_replicates, data, newdata, survey_weights, int_var, model_formulae, constituent_outcomes)

gam_rep_mn <- function(survey_design,
                        data,
                        newdata,
                        survey_weights,
                        model_formula,
                        constituent_outcomes,
                        survey_replicates,
                        k = 2,
                        int_var = NULL) {
  
  
  #Assign survey weights to the data.frame (so that gam finds it)
  data[, survey_weights := survey_weights ]
  #GAM object (to retrieve spline basis matrix)
  gam(
    data = data,
    family = multinom_wtd_mod(K = k),
    method = "REML",
    formula = model_formula,
    weights = survey_weights) -> gam_object
  
  
  #Get bootstrap estiamte of the variance
  withReplicates(design = survey_replicates,
                 theta = function(w, d) {
                   
                   #Reassign formulas to the current environment
                   f <- model_formula %>% lapply(function(form_part) 
                   {formula(form_part) -> form_part
                     environment(form_part) <- environment()
                     form_part
                   })
                   
                   
                   coef(gam(
                   family = multinom_wtd_mod(K = k),
                   method = "REML",
                   formula = f,
                   weights = w,
                   data = d)
                 )
                 }
  ) -> reps
  
  
  #Get length of coefficients and divide by k (to get N of coefficients per outcome)
  nc <- length(coef(reps))/k
  
  #Save coefficients in K - 1 lists
  coefs_list <- lapply(seq(0, k - 1, 1), function(i) coef(reps)[((i*nc) + 1):((i + 1)*nc)])
  
  #Obtain design matrix
  design_matrix <- predict(gam_object, newdata = newdata,
                           type = "lpmatrix")[, 1:nc]
  
  
  #Place the coefficients into an n by k matrix
  coef_matrix <- coef(reps) %>% matrix(ncol  = k)
  
  #Make predictions of prevalence from 1988 to 2023
  preds_dfs <- get_props_ci(coef_matrix = coef_matrix, vcov_matrix = vcov(reps), design_matrix = design_matrix,
                            predictors_df = newdata)
  
  #Get RRs comparing the last to first year
  contrasts_dfs <- get_rr_ci(coef_matrix = coef_matrix, vcov_matrix = vcov(reps), design_matrix = design_matrix[c(1, nrow(design_matrix)), ])

  #cats levels (reference, 1 to k, and sum of k)
  cats_levels <- c("== 0", paste0("== ", seq(1, k)), ">0")
  lapply(
    cats_levels, function(cat_level)
      
  #Get year-wise estimates (no model smoothing)
  svytotal_df_by(var = paste0("I(", as.character(model_formula[[1]])[2], cat_level, ")"), 
                 by_var = "year",
                 survey_object = survey_design) %>%
    filter(value == TRUE)
  
  ) -> year_dfs
  
  #Summarize output in a list
  lapply(1:(k + 2),
         function(i) 
           list("preds_df" = preds_dfs[[i]], 
                "year_df" = year_dfs[[i]], 
                "contrasts_df" = contrasts_dfs[[i]])
         ) -> output
  
  #Name output by constituent outcomes
  names(output) <- c("baseline", constituent_outcomes)
  
  #Return
  output
  
}

```

#### Interaction

```{r}

#Function to automate GAM estimation for risk factors underlying HF
#This function is used for multinomial outcomes (e.g., HTN stages and DM/PreDM)
#int_var is only defined here for the sake of compatibility with the subgroup-specific function because they will both be used in the same lapply loop
gam_rep_mn_int <- function(survey_design,
                           data,
                           newdata,
                           survey_weights,
                           model_formula,
                           constituent_outcomes,
                           k = 2,
                           survey_replicates,
                           int_var = NULL,
                           return_gam_object = FALSE) {
  
  
  #Assign survey weights to the data.frame (so that gam finds it)
  data[, survey_weights := survey_weights]
  #GAM object (to retrieve spline basis matrix)
  gam(
    data = data,
    family = multinom_wtd_mod(K = k),
    method = "REML",
    formula = model_formula,
    weights = survey_weights) -> gam_object
  
  
  #Get bootstrap estiamte of the variance
  withReplicates(design = survey_replicates,
                 theta = function(w, d) {
                   
                   #Reassign formulas to the current environment
                   f <- model_formula %>% lapply(function(form_part) 
                   {formula(form_part) -> form_part
                     environment(form_part) <- environment()
                     form_part
                   })
                   
                   
                   coef(gam(
                     family = multinom_wtd_mod(K = k),
                     method = "REML",
                     formula = f,
                     weights = w,
                     data = d)
                   )
                 }
  ) -> reps
  
  
  #Get length of coefficients and divide by k (to get N of coefficients per outcome)
  nc <- length(coef(reps))/k
  
  #Save coefficients in K - 1 lists
  coefs_list <- lapply(seq(0, k - 1, 1), function(i) coef(reps)[((i*nc) + 1):((i + 1)*nc)])
  
  #Obtain design matrix
  design_matrix <- predict(gam_object, newdata = newdata,
                           type = "lpmatrix")[, 1:nc]
  
  #Place the coefficients into an n by k matrix
  coef_matrix <- coef(reps) %>% matrix(ncol  = k)
  
  #Make predictions of prevalence from 1988 to 2023
  preds_dfs <- get_props_ci(coef_matrix = coef_matrix, 
                            vcov_matrix = vcov(reps), 
                            design_matrix = design_matrix,
                            predictors_df = newdata)
  
  #Get row numbers of which to get RRs based on
  data.table(newdata) -> newdata
  newdata[year == max(year) | year == min(year), which = T] %>% split(rep(1:(length(.)/2), each = 2)) -> i_list
  
  #Get RRs comparing the last to first year for each year
  lapply(i_list,
         function(i)
           get_rr_ci(coef_matrix = coef_matrix, vcov_matrix = vcov(reps), design_matrix = design_matrix[c(i[[1]], i[[2]]), ]) %>%
           rbindlist(idcol = "outcome") %>%
           mutate(int_factor = newdata[i[[1]],] %>% select(-year) %>% pull %>% unique)
         ) %>%
    rbindlist -> contrasts_df
  
  #Place RRs into separate lists for each outcome
  lapply(contrasts_df[, unique(outcome)],
         function(outcome_cat) contrasts_df[outcome == outcome_cat]) -> contrasts_dfs
  names(contrasts_dfs) <- names(preds_dfs)
  
  #Get number of unique factors to loop over
  newdata[, get(int_var) %>% unique] -> unique_factors
  
  #Get year-wise estimates (no model smoothing)
  #cats levels (reference, 1 to k, and sum of k)
  cats_levels <- c("== 0", paste0("== ", seq(1, k)), ">0")
  
  #Loop over k outcomes
  lapply(
    cats_levels, function(cat_level)
      
      #Get year-wise estimates (no model smoothing) per interaction level
      lapply(unique_factors,
             function(level) {
               
               #Subset expression
               subset_survey <- do.call("subset", 
                                        list(x = survey_design, 
                                             subset = parse(text = paste0(int_var, " == \"", level, "\""))
                                        )
               )
               
               #Get year-wise estimates (no model smoothing)
               svytotal_df_by(var = paste0("I(", as.character(model_formula[[1]])[2], cat_level, ")"), 
                              by_var = "year",
                              survey_object = subset_survey) %>%
                 filter(value == TRUE) %>%
                 mutate(int_factor = level) -> year_level_dfs
             }
      ) %>%
      rbindlist-> year_df
    
  ) -> year_dfs
  
  
  #Identify interaction indices (containing :)
  which(str_detect(names(gam_object$coefficients), ":")) -> selection_vector
  #P-value for interaction
  mgcv:::testStat(p = coef(reps)[selection_vector], V = vcov(reps)[selection_vector, selection_vector],
                  X = gam_object$R[, selection_vector],
                  rank = gam_object$edf1[selection_vector] %>% sum,
                  type = 0,
                  res.df = -1)$pval -> p_int
  
  
  #Return gam_object if true
if(!return_gam_object) {gam_object <- NULL}
  
  #Summarize output in a list
  lapply(1:(k + 2),
         function(i) 
           list("preds_df" = preds_dfs[[i]], 
                "year_df" = year_dfs[[i]], 
                "contrasts_df" = contrasts_dfs[[i]],
                "vcov" = vcov(reps),
                "coef" = coef(reps),
                "gam_object" = gam_object)
  ) -> output
  
  #Name output by constituent outcomes
  names(output) <- c("baseline", constituent_outcomes)
  
  output$p_interaction <- p_int
  
  #Return
  output
}


```



### Prediction dataframes

```{r}
#Overall
newdata_overall <- data.frame(year = 1988:2023)

#Sex
newdata_sex <- data.frame(year = rep(1988:2023, times = 2),
                     sex = rep(c("Women", "Men"), 
                               each = length(1988:2023)))
newdata_sex$sex_ord <- newdata_sex$sex

#Race/ethnicity
newdata_race <- data.frame(year = rep(1988:2023, times = 2),
                     race_binary = rep(c("Non-Hispanic White", "Non-Hispanic Black, Hispanic, or Other Minority"), 
                               each = length(1988:2023)))
newdata_race$race_binary_ord <- newdata_race$race_binary
```


### Setup

```{r}
#This block will set up a list that will be used recurrently throughout the next modelling tasks


overall_form <- .q(~ s(year, k = gam_k))
sex_form <- .q(~ sex + s(year, k = gam_k) + s(year, by = sex_ord, k = gam_k))
race_form <- .q(~ race_binary + s(year, k = gam_k) + s(year, by = race_binary_ord, k = gam_k))


#Analyses types
int_vars <- c("overall", "sex", "race_binary")
#
#Data list
newdata_list <- list(newdata_overall, newdata_sex, newdata_race)
#Functions list
gam_rep_list <- list(gam_rep_mn, gam_rep_mn_int, gam_rep_int)
#Create a base list (only missing different formulae)
list(trends_function = gam_rep_list,
     newdata = newdata_list,
     int_vars = int_vars) -> rfs_base_list 

```

## HF Prevalence

### Model

```{r}
#Set up formula
list(
  paste0("hf", c(overall_form)),
  paste0("hf", c(sex_form)),
  paste0("hf", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep_q",
                          "sdata_q", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = sdata_q,
              data = full_data,
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[, int_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = sdata_rep_q
            )
          }
) -> hf_trends
names(hf_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

### Figures

#### Create figure directories

```{r}

dir.create(path = paste0("Figures/", int_vars[1]), recursive = T)
dir.create(path = paste0("Figures/", int_vars[2]))
dir.create(path = paste0("Figures/", int_vars[3]))

```

#### Overall (crude proportions)

```{r}
trends_hf_fig <- ggplot(data = hf_trends$overall$preds_df,
      aes(x = year,
          y = pe,
          ymin = lci,
          ymax = uci)) +
  #Add point estimates based on cycle-specific estimates
  geom_pointrange(inherit.aes = F,
                  data =  hf_trends$overall$year_df %>%
                    filter(value == "Yes"),
                  aes(x = by_var,
                      y = prop_pe*100,
                      ymin = prop_lci*100,
                      ymax = prop_uci*100),
                  color = "grey10",
                  alpha = 0.4,
                  fatten = 4,
                  lwd = 1) +
  #Add Lines & Ribbons
  geom_ribbon(alpha = 0.5, lwd = 0, fill = "lightblue2") +
  geom_line(lwd = 1.5, color = "maroon") +
  geom_point(size = 3, color = "maroon") +
  #Add scales
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Crude prevalence (%)",
    breaks = seq(0, 100, 0.5)
  ) +
  #Add title
  ggtitle(paste0("Prevalence (proportional) of self-reported heart failure in the US, 1988-2023"),
          subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. The red line represents model-smoothed temporal trends and the blue shade represents its 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"))

ggsave(trends_hf_fig,
       filename = paste0("Figures/overall/HF Trends [Crude Proportions].pdf"),
     width = 16, height = 9,
     dpi = 600)
```


#### Overall (crude N)

##### Data

```{r}
full_data[age >= 20 & !is.na(hf), 
          .(N = sum(int_weight)), 
          by = survey_cycle][, .(survey_cycle, 
                                 first_year = word(survey_cycle, 1, sep = " to ") %>% as.numeric,
                                 last_year = word(survey_cycle, 2, sep = " to ") %>% as.numeric,
                                 N
                                 )][
                                   , .(year = seq(first_year, last_year), N = rep(N, last_year - first_year + 1)),
                                   by = survey_cycle
                                 ] %>%
  #One row will be repated twice (for 1991)
  unique(by = "year") -> n_by_year
merge(n_by_year,
      hf_trends$overall$preds_df,
      by = "year"
      ) -> n_props_df
n_props_df[, hf_n_pe := (pe/100)*N]
n_props_df[, hf_n_lci := (lci/100)*N]
n_props_df[, hf_n_uci := (uci/100)*N]

```

##### Plot

```{r}
trends_hf_n_fig <- ggplot(data = n_props_df,
      aes(x = year,
          y = hf_n_pe,
          ymin = hf_n_lci,
          ymax = hf_n_uci)) +
  #Add point estimates based on cycle-specific estimates
  geom_pointrange(inherit.aes = F,
                  data =  hf_trends$overall$year_df %>%
                    filter(value == "Yes"),
                  aes(x = by_var,
                      y = pe,
                      ymin = lci,
                      ymax = uci),
                  color = "grey10",
                  alpha = 0.4,
                  fatten = 4,
                  lwd = 1) +
  #Add Lines & Ribbons
  geom_ribbon(alpha = 0.5, lwd = 0, fill = "lightblue2") +
  geom_line(lwd = 1.5, color = "maroon") +
  geom_point(size = 3, color = "maroon") +
  #Add scales
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Crude prevalence (in millions)",
    labels = function(x) x/1000000,
    breaks = seq(0, 10000000, 1000000)
  ) +
  #Add title
  ggtitle(paste0("Prevalence (absolute number) of self-reported heart failure in the US, 1988-2023"),
          subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. The red line represents model-smoothed temporal trends and the blue shade represents its 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"))

ggsave(trends_hf_n_fig,
       filename = paste0("Figures/overall/HF Trends [Crude N].pdf"),
     width = 16, height = 9,
     dpi = 600)

```


#### Sex-stratified

```{r}
trends_sex_hf_fig <- ggplot(data = hf_trends$sex$preds_df %>% 
                              mutate(sex = str_replace_all(sex,
                                                           c("Men" = "Male", 
                                                             "Women" = "Female"))),
      aes(x = year,
          y = pe,
          color = sex,
          fill = sex,
          ymin = lci,
          ymax = uci)) +
  #Add point estimates based on cycle-specific estimates
  geom_pointrange(inherit.aes = F,
                  data =  hf_trends$sex$year_df %>%
                    filter(value == "Yes") %>% 
                              mutate(int_factor = str_replace_all(int_factor,
                                                           c("Men" = "Male", 
                                                             "Women" = "Female"))),
                  aes(x = by_var,
                      y = prop_pe*100,
                      ymin = prop_lci*100,
                      ymax = prop_uci*100,
                      color = int_factor),
                  alpha = 0.15,
                  fatten = 4,
                  lwd = 1) +
  #Add Lines & Ribbons
  geom_ribbon(alpha = 0.05, lwd = 0) +
  geom_line(lwd = 1.5) +
  geom_point(size = 4) +
  #Add scales
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Prevalence (%)",
    breaks = seq(0, 100, 0.5)
  ) +
  scale_fill_stata(name = NULL) +
  scale_color_stata(name = NULL) +
  #Add title
  ggtitle(paste0("Prevalence of self-reported heart failure in the US stratified by sex, 1988-2023"),
          subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. Solid lines represent model-smoothed temporal trends and the shade represents its 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        legend.key.width = unit(3, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.position = c(0.1, 0.825),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.text = element_text(size = 20, face = "bold"))

ggsave(trends_sex_hf_fig,
       filename = paste0("Figures/HF Trends - Sex Stratified.pdf"),
     width = 16, height = 9,
     dpi = 600)
```

#### Race/Ethnicity-stratified

```{r}
trends_race_hf_fig <- ggplot(data = hf_trends$race$preds_df,
      aes(x = year,
          y = pe,
          color = race_binary,
          fill = race_binary,
          ymin = lci,
          ymax = uci)) +
  #Add point estimates based on cycle-specific estimates
  geom_pointrange(inherit.aes = F,
                  data =  hf_trends$race$year_df %>%
                    filter(value == "Yes"),
                  aes(x = by_var,
                      y = prop_pe*100,
                      ymin = prop_lci*100,
                      ymax = prop_uci*100,
                      color = int_factor),
                  alpha = 0.15,
                  fatten = 4,
                  lwd = 1) +
  #Add Lines & Ribbons
  geom_ribbon(alpha = 0.05, lwd = 0) +
  geom_line(lwd = 1.5) +
  geom_point(size = 4) +
  #Add scales
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Prevalence (%)",
    breaks = seq(0, 100, 0.5)
  ) +
  scale_fill_stata(name = NULL) +
  scale_color_stata(name = NULL) +
  #Add title
  ggtitle(paste0("Prevalence of self-reported heart failure in the US stratified by race/ethnicity, 1988-2023"),
          subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. Solid lines represent model-smoothed temporal trends and the shade represents its 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        legend.key.width = unit(3, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.position = c(0.3, 0.825),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.text = element_text(size = 20, face = "bold"))

ggsave(trends_race_hf_fig,
       filename = paste0("Figures/HF Trends - Race-Ethnicity Stratified.pdf"),
     width = 16, height = 9,
     dpi = 600)
```

## Trends in RFs

### Multinomial

#### Diabetes

```{r}
#Set up formula
list(
  list(
    paste0("dm_cats", c(overall_form)),
    overall_form
  ),
  list(
    paste0("dm_cats", c(sex_form)),
    sex_form
  ),
  list(
    paste0("dm_cats", c(race_form)),
    race_form
  )
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep_mn, gam_rep_mn_int, gam_rep_mn_int)
#Multinomial outcomes
rfs_base_list$constituent_outcomes <- c("pre_dm", "dm", "pre_dm_or_dm")

#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
       function(i)
       {
         #Use self-made function
         rfs_base_list$trends_function[[i]](
           survey_design = subset(sdata, hf == "Yes"),
           data = full_data[hf == "Yes"],
           newdata = rfs_base_list$newdata[[i]],
           survey_weights = full_data[hf == "Yes", exam_weight],
           model_formula =  rfs_base_list$model_formulae[[i]] %>% lapply(formula),
           int_var = rfs_base_list$int_vars[[i]],
           survey_replicates = subset(sdata_rep, hf == "Yes"),
           constituent_outcomes = rfs_base_list$constituent_outcomes
         )
       }
) -> dm_trends
names(dm_trends) <- int_vars
stopCluster(cl)


```

#### Obesity

```{r}
#Set up formula
list(
  list(
    paste0("obesity_cats", c(overall_form)),
    overall_form
  ),
  list(
    paste0("obesity_cats", c(sex_form)),
    sex_form
  ),
  list(
    paste0("obesity_cats", c(race_form)),
    race_form
  )
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep_mn, gam_rep_mn_int, gam_rep_mn_int)
#Multinomial outcomes
rfs_base_list$constituent_outcomes <- c("obesity12", "obesity3", "obesity")


#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep", 
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
       function(i)
       {
         #Use self-made function
         rfs_base_list$trends_function[[i]](
           survey_design = subset(sdata, hf == "Yes"),
           data = full_data[hf == "Yes"],
           newdata = rfs_base_list$newdata[[i]],
           survey_weights = full_data[hf == "Yes", exam_weight],
           model_formula =  rfs_base_list$model_formulae[[i]] %>% lapply(formula),
           int_var = rfs_base_list$int_vars[[i]],
           survey_replicates = subset(sdata_rep, hf == "Yes"),
           constituent_outcomes = rfs_base_list$constituent_outcomes
         )
       }
) -> obesity_trends
names(obesity_trends) <- int_vars
stopCluster(cl)
Sys.time() - start_time

```

#### HTN-related

##### Elevated BP

```{r}
#Set up formula
list(
  list(
    paste0("htn_cats", c(overall_form)),
    overall_form
  ),
  list(
    paste0("htn_cats", c(sex_form)),
    sex_form
  ),
  list(
    paste0("htn_cats", c(race_form)),
    race_form
  )
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep_mn, gam_rep_mn_int, gam_rep_mn_int)
#Multinomial outcomes
rfs_base_list$constituent_outcomes <- c("htn1", "htn2", "htn")


#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes", exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% lapply(formula),
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes"),
              constituent_outcomes = rfs_base_list$constituent_outcomes
            )
          }
) -> htn_trends
names(htn_trends) <- int_vars
stopCluster(cl)
Sys.time() - start_time 

```

##### Elevated BP among HTN history

```{r}

#Set up formula
list(
  list(
    paste0("htn_cats", c(overall_form)),
    overall_form
  ),
  list(
    paste0("htn_cats", c(sex_form)),
    sex_form
  ),
  list(
    paste0("htn_cats", c(race_form)),
    race_form
  )
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep_mn, gam_rep_mn_int, gam_rep_mn_int)
#Multinomial outcomes
rfs_base_list$constituent_outcomes <- c("htn1", "htn2", "htn")


#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes" & htn_hist == "Yes"),
              data = full_data[hf == "Yes" & htn_hist == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes" & htn_hist == "Yes", exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% lapply(formula),
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes" & htn_hist == "Yes"),
              constituent_outcomes = rfs_base_list$constituent_outcomes
            )
          }
) -> htn_pos_hist_trends
names(htn_pos_hist_trends) <- int_vars
stopCluster(cl)
Sys.time() - start_time 

```

### Binomial

####Myocardial Infarction

```{r}
#### Myocardial Infarction
#Set up formula
list(
    paste0("mi", c(overall_form)),
    paste0("mi", c(sex_form)),
    paste0("mi", c(race_form))
  ) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep_q",
                          "sdata_q", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata_q, hf == "Yes"),
              data = full_data[hf == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes", int_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep_q, hf == "Yes")
            )
          }
) -> mi_trends
names(mi_trends) <- int_vars
stopCluster(cl)
Sys.time() - start_time 
```

#### Lipid-related

##### Hypercholesterolemia

```{r}
#Set up formula
list(
  paste0("hypercholesterol", c(overall_form)),
  paste0("hypercholesterol", c(sex_form)),
  paste0("hypercholesterol", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (binomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL


#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes", exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> hypercholesterol_trends
names(hypercholesterol_trends) <- int_vars
stopCluster(cl)
Sys.time() - start_time 

```


##### LDL

```{r}

#Set up formula
list(
  paste0("hyperldl", c(overall_form)),
  paste0("hyperldl", c(sex_form)),
  paste0("hyperldl", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 3
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep_fast",
                          "sdata_fast", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata_fast, hf == "Yes"),
              data = full_data[hf == "Yes" & year <= 2020],
              newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
              survey_weights = full_data[hf == "Yes" & year <= 2020, fast_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep_fast, hf == "Yes")
            )
          }
) -> hyperldl_trends
names(hyperldl_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

##### Triglycerides

```{r}

#Set up formula
list(
  paste0("hypertg", c(overall_form)),
  paste0("hypertg", c(sex_form)),
  paste0("hypertg", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL
#Filter to 2020 or before 
lapply(1:3, function(i) rfs_base_list$newdata[[i]] %>% filter(year <= 2020)) -> rfs_base_list$newdata


#Set up basis functions
gam_k <- 3
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep_fast",
                          "sdata_fast", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata_fast, hf == "Yes"),
              data = full_data[hf == "Yes" & year <= 2020],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes" & year <= 2020, fast_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep_fast, hf == "Yes")
            )
          }
) -> hypertg_trends
names(hypertg_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

#####HC history

```{r}

#Set up formula
list(
  paste0("dyslipid_hist2", c(overall_form)),
  paste0("dyslipid_hist2", c(sex_form)),
  paste0("dyslipid_hist2", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes", exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> dyslipid_hist2_trends
names(dyslipid_hist2_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

#####Statin
```{r}
#Statins
#Set up formula
list(
  paste0("statin", c(overall_form)),
  paste0("statin", c(sex_form)),
  paste0("statin", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
i = 1
gam_k <- 3
#Use self-made function
rfs_base_list$trends_function[[i]](
  survey_design = subset(sdata, hf == "Yes" & dyslipid_hist2 == "Yes"),
  data = full_data[hf == "Yes"  & dyslipid_hist2 == "Yes" & year <= 2020],
  newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
  survey_weights = full_data[hf == "Yes"  & dyslipid_hist2 == "Yes" & year <= 2020, exam_weight],
  model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
  int_var = rfs_base_list$int_vars[[i]],
  survey_replicates = subset(sdata_rep, hf == "Yes" & dyslipid_hist2 == "Yes")) -> statin_trends

```

#### HTN-related

##### HTN history

```{r}

#Set up formula
list(
  paste0("htn_hist2", c(overall_form)),
  paste0("htn_hist2", c(sex_form)),
  paste0("htn_hist2", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 4
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes"],
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[hf == "Yes", exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> htn_hist2_trends
names(htn_hist2_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()
```

##### HTN Meds 

```{r}

#Use fewer knots
gam_k <- 3
#CCB
#Set up formula
list(
  paste0("ccb", c(overall_form)),
  paste0("ccb", c(sex_form)),
  paste0("ccb", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
i = 1
#Use self-made function
rfs_base_list$trends_function[[i]](
  survey_design = subset(sdata, hf == "Yes" & htn_hist2 == "Yes"),
  data = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020],
  newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
  survey_weights = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020, exam_weight],
  model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
  int_var = rfs_base_list$int_vars[[i]],
  survey_replicates = subset(sdata_rep, hf == "Yes" & htn_hist2 == "Yes")) -> ccb_trends

#ACEI/ARB
#Set up formula
list(
  paste0("acei_arb", c(overall_form)),
  paste0("acei_arb", c(sex_form)),
  paste0("acei_arb", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
i = 1
#Use self-made function
rfs_base_list$trends_function[[i]](
  survey_design = subset(sdata, hf == "Yes" & htn_hist2 == "Yes"),
  data = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020],
  newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
  survey_weights = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020, exam_weight],
  model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
  int_var = rfs_base_list$int_vars[[i]],
  survey_replicates = subset(sdata_rep, hf == "Yes" & htn_hist2 == "Yes")) -> acei_arb_trends

#BB
#Set up formula
list(
  paste0("bb", c(overall_form)),
  paste0("bb", c(sex_form)),
  paste0("bb", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
i = 1
#Use self-made function
rfs_base_list$trends_function[[i]](
  survey_design = subset(sdata, hf == "Yes" & htn_hist2 == "Yes"),
  data = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020],
  newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
  survey_weights = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020, exam_weight],
  model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
  int_var = rfs_base_list$int_vars[[i]],
  survey_replicates = subset(sdata_rep, hf == "Yes" & htn_hist2 == "Yes")) -> bb_trends


#Thiazides
#Set up formula
list(
  paste0("thiazide", c(overall_form)),
  paste0("thiazide", c(sex_form)),
  paste0("thiazide", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
i = 1
#Use self-made function
rfs_base_list$trends_function[[i]](
  survey_design = subset(sdata, hf == "Yes" & htn_hist2 == "Yes"),
  data = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020],
  newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
  survey_weights = full_data[hf == "Yes"  & htn_hist2 == "Yes" & year <= 2020, exam_weight],
  model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
  int_var = rfs_base_list$int_vars[[i]],
  survey_replicates = subset(sdata_rep, hf == "Yes" & htn_hist2 == "Yes")) -> thiazide_trends



```


###CKD-related

#### CKD345/albuminuria 

```{r}
#Set up formula
list(
  paste0("ckd345_albumin", c(overall_form)),
  paste0("ckd345_albumin", c(sex_form)),
  paste0("ckd345_albumin", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 3
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes" & year <= 2020],
              newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
              survey_weights = full_data[hf == "Yes" & year <= 2020, exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> ckd345_albumin_trends
names(ckd345_albumin_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

#### CKD345 

```{r}
#Set up formula
list(
  paste0("ckd345", c(overall_form)),
  paste0("ckd345", c(sex_form)),
  paste0("ckd345", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 3
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes" & year <= 2020],
              newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
              survey_weights = full_data[hf == "Yes" & year <= 2020, exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> ckd345_trends
names(ckd345_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

#### Albuminuria 

```{r}
#Set up formula
list(
  paste0("albuminuria", c(overall_form)),
  paste0("albuminuria", c(sex_form)),
  paste0("albuminuria", c(race_form))
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL

#Set up basis functions
gam_k <- 3
#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(3)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep",
                          "sdata", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1:3,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = subset(sdata, hf == "Yes"),
              data = full_data[hf == "Yes" & year <= 2020],
              newdata = rfs_base_list$newdata[[i]] %>% filter(year <= 2020),
              survey_weights = full_data[hf == "Yes" & year <= 2020, exam_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = subset(sdata_rep, hf == "Yes")
            )
          }
) -> albuminuria_trends
names(albuminuria_trends) <- int_vars
stopCluster(cl)
start_time - Sys.time()

```

###Age-standardized trends

#### Model

```{r}


#Set up basis functions
gam_k <- 4
#Set up formula
list(
  paste0("hf", "~ s(year, k = 4) + ", "s(age, k = 4) + ", "ti(year, age, k = 3, bs = \"tp\")")
) -> model_formulae
rfs_base_list$model_formulae <- model_formulae
#Trend functions (multinomial)
rfs_base_list$trends_function <- list(gam_rep, gam_rep_int, gam_rep_int)
rfs_base_list$constituent_outcomes <- NULL
rfs_base_list$newdata[[1]] <- expand.grid(age = seq(20, 80, 1), year = seq(1988, 2023, 1))

#Start timer
start_time <- Sys.time()
#Adjust as necessary
cl <- makeCluster(1)
clusterExport(cl=cl, list("rfs_base_list", "sdata_rep_q",
                          "sdata_q", "gam_k", "full_data", 
                          "multinom_wtd_mod", "get_props_ci", "get_rr_ci", 
                          "svytotal_df", "svytotal_df_by", "r1", "r2",
                          "softmax"))
clusterEvalQ(cl, lapply(c("mgcv", "data.table", "dplyr", "Hmisc", "stringr", 
                          "survey", "svrep", "msm"), library, character.only = TRUE))
parLapply(cl,
          1,
          function(i)
          {
            #Use self-made function
            rfs_base_list$trends_function[[i]](
              survey_design = sdata_q,
              data = full_data,
              newdata = rfs_base_list$newdata[[i]],
              survey_weights = full_data[, int_weight],
              model_formula =  rfs_base_list$model_formulae[[i]] %>% formula,
              int_var = rfs_base_list$int_vars[[i]],
              survey_replicates = sdata_rep_q,
              return_gam_object = T
            )
          }
) -> hf_trends_aa
start_time - Sys.time()
stopCluster(cl)
names(hf_trends_aa) <- "overall"




```


#### Temporal trends (1988-2023)

##### Calculate

```{r}
#Obtain design matrix for desired predictions
predict(hf_trends_aa$overall$gam_object,
        newdata = rfs_base_list$newdata[[1]],
        type = "lpmatrix") -> d_matrix

#Get PEs
d_matrix %*% hf_trends_aa$overall$coef -> logit_pe
#Get vcov matrix
d_matrix %*% hf_trends_aa$overall$vcov %*% t(d_matrix) -> logit_vcov

#Get age-standardized weights
full_data[survey_cycle == "2021 to 2023", .(weight = sum(int_weight)), by = age][
  order(age), .(age, weight = weight/sum(weight))] -> std_weights
#Length of vector for age-standardized weights
weights_length <- std_weights[, length(weight)]

#Loop over from 1988 to 2023
lapply(
  1:length(1988:2023), #Loop over years,
  function(i) {
    
    #Get indidces for first and last age row in a given year
    index1 <- (((i - 1) * weights_length) + 1)
    index2 <- (i*weights_length)
    
    #Get probabilities across all ages for a given year
    p <- plogis(logit_pe[index1:index2]) 
    
    #Get variance/derivative of probabilities for each age
    var_p <- p * (1 - p)
    #Weight variance of probabilities for each age by the respective weight
    w_var_p <- var_p * std_weights[, weight]
    #Get vcov matrix for age-wise predictions in a given year
    vcov_logit <- logit_vcov[index1:index2, index1:index2]
    #Get variance for the age-standardized prediction
    var_std_p <- t(w_var_p) %*% vcov_logit %*% w_var_p
    #Get SD
    se_std_p <- sqrt(var_std_p)
    #Get point estimate
    std_p <- t(p) %*% std_weights[, weight]
    #Get LCI and UCI plus the PE
    std_ps <- as.numeric(std_p) + (c(-1.96, 0, 1.96) * as.numeric(se_std_p))
    #Combine in a data.table
    data.table(
      lci = std_ps[1],
      pe = std_ps[2],
      uci = std_ps[3]
    )
  }
) %>%
  rbindlist -> aa_preds_df
aa_preds_df[, year := 1988:2023]
aa_preds_df[, (c("lci", "pe", "uci")) := lapply(.SD, function(x) x * 100),
            .SDcols = c("lci", "pe", "uci")]
aa_preds_df[, pe_ci := paste0(r1(pe), " (", r1(lci), " to ", r1(uci), ")")]

```

##### Plot

```{r}
ggplot(data = aa_preds_df,
       aes(x = year, y = pe, ymin = lci, ymax = uci)
       ) +
    #Add Lines & Ribbons
  geom_ribbon(alpha = 0.5, lwd = 0, fill = "lightblue2") +
  geom_line(lwd = 1.5, color = "maroon") +
  geom_point(size = 3, color = "maroon") +
  #Add scales
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Age-standardized prevalence (%)",
    breaks = seq(0, 100, 0.2)
  ) +
  #Add title
  ggtitle(paste0("Age-standardized prevalence of self-reported heart failure in the US, 1988-2023"),
          subtitle = "The red line represents the temporal trend and the blue shade represents its 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 20, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        legend.key.width = unit(3, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.position = c(0.1, 0.825),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.text = element_text(size = 20, face = "bold")) -> trends_hf_aa_fig

ggsave(trends_hf_aa_fig,
       filename = paste0("Figures/Overall/HF Trends [Age-adjusted].pdf"),
     width = 16, height = 9,
     dpi = 600)


```


##### Risk ratio

```{r}

#Get indices for first and last age-group in 2023
i = 1
index1 <- (((i - 1) * weights_length) + 1)
index2 <- (i*weights_length)

#Get indices for first and last age-group in 2023
i = 36
index3 <- (((i - 1) * weights_length) + 1)
index4 <- (i*weights_length)


#Get probabilities for 1988 and 2023
p <- plogis(logit_pe[c(index1:index2, index3:index4)])

#Get variance
var_p <- p * (1 - p)
var_p <- diag(var_p)

#Get VCOV of logits for 1988/2023
vcov_logit <- logit_vcov[c(index1:index2, index3:index4), c(index1:index2, index3:index4)]
vcov_p <- var_p %*% vcov_logit %*% var_p

#Paste an expression together
xs <- paste0("x", 1:length(p))
# Expressions for each group
p[1:(length(p)/2)] -> p1
p[(1 + length(p)/2):(length(p))] -> p2
expr_p1 <- paste0(std_weights[, weight], "*x", c(1:length(p1)), collapse = "+")
expr_p2 <- paste0(std_weights[, weight], "*x", c((1 + length(p1)):(2*length(p2))), collapse = "+")
full_expr <- paste0("log((", expr_p1, ")", "/", "(", expr_p2, "))")

# Full log risk ratio expression
full_expr <- paste0("log((", expr_p1, ")/(", expr_p2, "))")

# Apply delta method
deltamethod(g = parse(text = full_expr),
            mean = c(p1, p2),
            cov = vcov_p) -> logrr_se

#Get LCI/PE/UCI
logrr <- log(sum(p2 * std_weights[, weight])/sum(p1 * std_weights[, weight])) + c(-1.96, 0, 1.96)*logrr_se
hf_trends_aa$overall$rr <- data.table(
  lci = exp(logrr[1]),
  pe = exp(logrr[2]),
  uci = exp(logrr[3])
)
hf_trends_aa$overall$rr[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]

```

####Age-wise prevalence of HF

#####Calculate

```{r}
#Number of unique age-groups
rfs_base_list$newdata[[1]] %>% pull(age) %>% uniqueN -> age_n

#Vector to select relevant elements (all ages in 1998 and 2023)
selection_vector <- c(1:age_n, (length(logit_pe) - age_n + 1):length(logit_pe))

#Get logit PEs and VCOV
logit_pe[selection_vector] -> logit_pe_age
logit_vcov[selection_vector, selection_vector] -> vcov_logit_age
logit_se_age <- diag(vcov_logit_age) %>% sqrt

#Put everything into a data.table
hf_age_prev_df <- data.table(
  pe = logit_pe_age %>% plogis,
  lci = (logit_pe_age - 1.96 * logit_se_age) %>% plogis,
  uci = (logit_pe_age + 1.96 * logit_se_age) %>% plogis
  )
hf_age_prev_df[, age := rfs_base_list$newdata[[1]] %>% pull(age) %>% unique %>% rep(times = 2)]
hf_age_prev_df[, year := c(1988, 2023) %>% rep(each = nrow(hf_age_prev_df)/2)]

#Round
hf_age_prev_df[, (c("lci", "pe", "uci")) := lapply(.SD, function(x) x * 100),
            .SDcols = c("lci", "pe", "uci")]
hf_age_prev_df[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]


```

#####Plot
```{r}
ggplot(data = hf_age_prev_df,
       aes(x = age, y = pe, ymin = lci, ymax = uci, 
           color = factor(year), fill = factor(year))
       ) +
    #Add Lines & Ribbons
  geom_ribbon(alpha = 0.1, lwd = 0) +
  geom_line(lwd = 1.5) +
  #Add scales
  scale_x_continuous(
    name = "Age",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Age-stratified Prevalence (%)",
    breaks = seq(0, 100, 1)
  ) +
  scale_color_jama(name = "Year") +
  scale_fill_jama(name = "Year") +
  #Add title
  ggtitle(paste0("Prevalence of self-reported heart failure in the US across the spectrum of age, 1988-2023"),
          subtitle = "Solid lines represents the point estimates and the shades represents the 95% CI.") +
  #Theme
  theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        legend.key.width = unit(3, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.position = c(0.1, 0.825),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.text = element_text(size = 20, face = "bold")) -> hf_age_strat_fig

ggsave(hf_age_strat_fig,
       filename = paste0("Figures/Overall/HF Trends [Age-stratified].pdf"),
     width = 16, height = 9,
     dpi = 600)

```


### REV

```{r}
#Predictions
full_data[, aa_preds := predict(hf_trends_aa$overall$gam_object, 
                                full_data[, .(age, year)],
                                type = "lpmatrix") %*% (hf_trends_aa$overall$coef)]

#REV due to year
gam(data = full_data,
   weights = int_weight,
   aa_preds ~ s(age, k = 4)) %>%
  summary %>%
  {1 - .$dev.expl} -> year_rexvar

#REV due to age
gam(data = full_data,
    weights = int_weight,
    aa_preds ~ s(year, k = 4)) %>%
  summary %>%
  {1 - .$dev.expl} -> age_rexvar


```

### RF Figures

#### Line graphs

```{r}
#Outer loop (over overall/race/sex)
lapply(int_vars,
       function(int_var) {

list("albuminuria" = albuminuria_trends[[int_var]], 
     "a reduced eGFR" = ckd345_trends[[int_var]],
     "chronic kidney disease" = ckd345_albumin_trends[[int_var]],
     "diabetes" = dm_trends[[int_var]]$dm,
     "prediabetes" = dm_trends[[int_var]]$pre_dm,
     "diabetes or pre-diabetes" = dm_trends[[int_var]]$pre_dm_or_dm,
     "obesity" = obesity_trends[[int_var]]$obesity,
     "class I or II obesity" = obesity_trends[[int_var]]$obesity12,
     "class III obesity" = obesity_trends[[int_var]]$obesity3,
     "elevated blood pressure" = htn_trends[[int_var]]$htn,
     "stage 1 hypertension" = htn_trends[[int_var]]$htn1,
     "stage 2 hypertension" = htn_trends[[int_var]]$htn2,
     "a history of hypertension" = htn_hist2_trends[[int_var]],
     "hypercholesterolemia" = hypercholesterol_trends[[int_var]], 
     "a history of hypercholesterolemia" = dyslipid_hist2_trends[[int_var]], 
     "elevated low-density lipoprotein cholesterol" = hyperldl_trends[[int_var]],
     "hypertriglyceridemia" = hypertg_trends[[int_var]],
     "a previous myocardial infarction" = mi_trends[[int_var]]) -> list_for_rf_figs


#Inner loop (over outcomes)
#Loop over the different items
lapply(1:length(list_for_rf_figs),
       function(i) {
         
       #Get plot label
       plot_label <- names(list_for_rf_figs[i])
       
       #Get list object by index
       list_object <- list_for_rf_figs[[i]]
       
       if(int_var == "sex") {
         list_object$preds_df[, sex := str_replace(sex, "Men", "Males")]
         list_object$preds_df[, sex := str_replace(sex, "Women", "Females")]
         list_object$year_df[, int_factor := str_replace(int_factor, "Men", "Males")]
         list_object$year_df[, int_factor := str_replace(int_factor, "Women", "Females")]
         }
       
       
       main_layer <- if(int_var == "overall") {
         ggplot(data = list_object$preds_df,
                aes(x = year,
                    y = pe,
                    ymin = lci,
                    ymax = uci)) +
           #Add point estimates based on cycle-specific estimates
           geom_pointrange(inherit.aes = F,
                           data = list_object$year_df,
                           aes(x = by_var,
                               y = prop_pe*100,
                               ymin = prop_lci*100,
                               ymax = prop_uci*100),
                           color = "grey10",
                           alpha = 0.4,
                           fatten = 4,
                           lwd = 1) +
           #Add Lines & Ribbons
           geom_ribbon(alpha = 0.5, lwd = 0, fill = "lightblue2") +
           geom_line(lwd = 1.5, color = "maroon") +
           geom_point(size = 3, color = "maroon") +
           
           #Add text annotation
           annotate(geom = "text",
                    size = 10*0.5,
                    fontface = "bold",
                    x = -Inf, y = Inf,
                    hjust = -0.1,
                    vjust = 1.4,
                    label = paste0(
                      list_object$preds_df[1, pe %>% round(1) %>% format(nsmall = 1)],
                      "% in 1988 to ",
                      list_object$preds_df[nrow(list_object$preds_df), pe %>% round(1) %>% format(nsmall = 1)], 
                      "% in ", list_object$preds_df[nrow(list_object$preds_df), year]
                    )
           )
       } else {
         ggplot(data = list_object$preds_df,
                aes(x = year,
                    y = pe,
                    ymin = lci,
                    ymax = uci,
                    color = !!as.symbol(int_var),
                    fill = !!as.symbol(int_var)
                    )) +
           #Add point estimates based on cycle-specific estimates
           geom_pointrange(inherit.aes = F,
                           data = list_object$year_df,
                           aes(x = by_var,
                               y = prop_pe*100,
                               color = int_factor,
                               ymin = prop_lci*100,
                               ymax = prop_uci*100),
                           alpha = 0.4,
                           fatten = 4,
                           lwd = 1) +
           #Add Lines & Ribbons
           geom_ribbon(alpha = 0.1, color = "transparent", lwd = 0) +
           geom_line(lwd = 1.5) +
           geom_point(size = 3) +
           scale_color_jama(name = NULL) +
           scale_fill_jama(name = NULL)
       }
       
       
  #Plot! 
  trends_fig <- main_layer +
    #Add scales
    scale_x_continuous(
      name = "Year",
      breaks = seq(0, 5000, 2),
      expand = c(0.02, 0.0)
    ) +
    scale_y_continuous(
      name = "Prevalence (%)",
      breaks = seq(0, 100, 5)
    ) +
    #Add title
    ggtitle(paste0("Prevalence of ", plot_label, " among patients with heart failure in the US, 1988-", list_object$preds_df[, max(year)]),
            subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. The red line represents model-smoothed temporal trends and the blue shade represents its 95% CI.") +
    #Theme
    theme_bw() +
    theme(text = element_text(size = 23),
          plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
          plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
          axis.text.x = element_text(size = 20, face = "bold"),
          axis.text.y = element_text(size = 20, face = "bold"),
          axis.title.x = element_text(size = 25, face = "bold"),
          axis.title.y = element_text(size = 25, face = "bold"),
          axis.line = element_line(colour = "black", linewidth = 1.2),
          plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
          legend.key.width = unit(3, "cm"),
          legend.key.height = unit(1, "cm"),
          legend.position = "top",
          legend.background = element_rect(fill = "transparent"),
          legend.key = element_rect(fill = "transparent"),
          legend.title = element_text(size = 20, face = "bold"),
          legend.text = element_text(size = 20, face = "bold"))
  
  #Save
  ggsave(trends_fig,
         filename = paste0("Figures/", int_var, "/Trends in ", plot_label, " [", int_var, "].pdf"),
         width = 16, height = 9,
         dpi = 600)
  
  #Save object in list
  trends_fig
}) -> trends_fig_rfs

#Name list
names(trends_fig_rfs) <- names(list_for_rf_figs)

#Return
trends_fig_rfs

       }


) -> trends_fig_rfs

names(trends_fig_rfs) <- int_vars

```


#### Cumulative figures

```{r}


list(
  "insulin resistance" = list(
    "Diabetes" = dm_trends$overall$dm$preds_df,
    "Prediabetes" = dm_trends$overall$pre_dm$preds_df,
    "No diabetes or prediabetes" = dm_trends$overall$baseline$preds_df
     ),
  "obesity" = list(
    "Class III obesity" = obesity_trends$overall$obesity3$preds_df,
    "Class I or II obesity" = obesity_trends$overall$obesity12$preds_df,
    "No obesity" = obesity_trends$overall$baseline$preds_df),
    "elevated blood pressure" = list(
      "Stage 1 hypertension" = htn_trends$overall$htn1$preds_df,
      "Stage 2 hypertension" = htn_trends$overall$htn2$preds_df,
      "Normal blood pressure" = htn_trends$overall$baseline$preds_df
  )
  ) -> list_for_cum_figs

lapply(1:length(list_for_cum_figs),
       function(i) {
         
         
         #label
         label <- names(list_for_cum_figs)[i]
         
         #Save factor order in the same way lists are ordered
         factor_order <- names(list_for_cum_figs[[i]])
         #Name outcome categories
         lapply(1:length(list_for_cum_figs[[i]]),
                function(i2)
                list_for_cum_figs[[i]][[i2]][, cat := names(list_for_cum_figs[[i]])[[i2]]]
                ) -> list_for_cum_figs[[i]]
         
         
         #Bind list components into a single data.table
         cum_fig_df <- list_for_cum_figs[[i]] %>% rbindlist
         cum_fig_df[, cat := factor(cat, factor_order)]
         
         
         ggplot(data = cum_fig_df,
                aes(x = year,
                    y = pe,
                    fill = cat)) +
           geom_area(lwd = 1, color = "black") +
           ggtitle(paste0("Changes in the prevalence of ", label, " among patients with heart failure in the US, 1988 to 2023")) +
           scale_y_continuous(expand = c(0, 0),
                              name = "Prevalence (%)",
                              limits = c(0, 101),
                              breaks = seq(10, 100, 10),
                              sec.axis = dup_axis(name = "Prevalence (%)")
                              ) +
           scale_x_continuous(expand = c(0, 0),
                              name = "Year",
                              breaks = seq(0, 5000, 2),
                              limits = c(1988, 2023)) +
           scale_fill_jama(name = NULL) +
           theme_bw() +
           theme(text = element_text(size = 23, face = "bold"),
                 plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
                 axis.text.x = element_text(size = 20, face = "bold"),
                 axis.text.y = element_text(size = 20, face = "bold"),
                 axis.title.x = element_text(size = 25, face = "bold"),
                 axis.title.y = element_text(size = 25, face = "bold"),
                 axis.line = element_line(colour = "black", size = 0),
          plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
                 panel.grid = element_blank(),
                 legend.position = "bottom",
                 legend.key.width = unit(4, "cm"),
                 legend.key.height = unit(0.8, "cm")
                 ) -> cum_fig
              
         #Save
         ggsave(cum_fig,
                filename = paste0("Figures/overall/Cumulative figure for ", label, ".pdf"),
                dpi = 600,
                width = 16, height = 9)
         
         #Return
         cum_fig
       } 
       ) -> cum_figs

names(cum_figs) <- names(list_for_cum_figs)


```


### Med figures

```{r}



#Create list containing medication trends
list("use of ACEis or ARBs" = acei_arb_trends, 
    "use of beta-blockers" = bb_trends, 
    "use of calcium channel blockers" = ccb_trends, 
    "use of thiazides" = thiazide_trends, 
    "use of statins" = statin_trends
    ) -> list_for_med_figs


#Inner loop (over outcomes)
#Loop over the different items
lapply(1:length(list_for_med_figs),
      function(i) {
        
        #Get plot label
        plot_label <- names(list_for_med_figs[i])
        
        #Get list object by index
        list_object <- list_for_med_figs[[i]]
        
        
        #Plot! 
        trends_fig <- ggplot(data = list_object$preds_df,
                             aes(x = year,
                                 y = pe,
                                 ymin = lci,
                                 ymax = uci)) +
          #Add point estimates based on cycle-specific estimates
          geom_pointrange(inherit.aes = F,
                          data = list_object$year_df,
                          aes(x = by_var,
                              y = prop_pe*100,
                              ymin = prop_lci*100,
                              ymax = prop_uci*100),
                          color = "grey10",
                          alpha = 0.4,
                          fatten = 4,
                          lwd = 1) +
          #Add Lines & Ribbons
          geom_ribbon(alpha = 0.5, lwd = 0, fill = "lightblue2") +
          geom_line(lwd = 1.5, color = "maroon") +
          geom_point(size = 4, color = "maroon") +
          
          #Add text annotation
          annotate(geom = "text",
                   size = 10*0.5,
                   fontface = "bold",
                   x = -Inf, y = Inf,
                   hjust = -0.1,
                   vjust = 1.4,
                   label = paste0(
                     list_object$preds_df[1, pe %>% round(1) %>% format(nsmall = 1)],
                     "% in 1988 to ",
                     list_object$preds_df[nrow(list_object$preds_df), pe %>% round(1) %>% format(nsmall = 1)], 
                     "% in ", list_object$preds_df[nrow(list_object$preds_df), year]
                   )
          ) +
          #Add scales
          scale_x_continuous(
            name = "Year",
            breaks = seq(0, 5000, 2),
            expand = c(0.02, 0.0)
          ) +
          scale_y_continuous(
            name = "Prevalence (%)",
            breaks = seq(0, 100, 5)
          ) +
          #Add title
          ggtitle(paste0("Prevalence of ", plot_label, " among patients with heart failure in the US, 1988-2020"),
                  subtitle = "Dots and whiskers represent point estimates and 95% CI for each survey cycle. The red line represents model-smoothed temporal trends and the blue shade represents its 95% CI.") +
          #Theme
          theme_bw() +
          theme(text = element_text(size = 23),
                plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
                plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
                axis.text.x = element_text(size = 20, face = "bold"),
                axis.text.y = element_text(size = 20, face = "bold"),
                axis.title.x = element_text(size = 25, face = "bold"),
                axis.title.y = element_text(size = 25, face = "bold"),
                axis.line = element_line(colour = "black", linewidth = 1.2),
                plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
                legend.key.width = unit(3, "cm"),
                legend.key.height = unit(1, "cm"),
                legend.position = "top",
                legend.background = element_rect(fill = "transparent"),
                legend.key = element_rect(fill = "transparent"),
                legend.title = element_text(size = 20, face = "bold"),
                legend.text = element_text(size = 20, face = "bold"))
        
        #Save
        ggsave(trends_fig,
               filename = paste0("Figures/overall", "/Trends in ", plot_label, ".pdf"),
               width = 16, height = 9,
               dpi = 600)
        
        #Save object in list
        trends_fig
      }) -> trends_fig_meds

#Name list
names(trends_fig_meds) <- names(list_for_med_figs)




```

##Collate multipanel figures

###Rescale

```{r}
rescale_ggplot <-
  theme(
  plot.title = element_text(size = 7, face = "bold", color = "transparent", hjust = 0),
  plot.subtitle = element_blank(),
  axis.text.x = element_text(size = rel(0.55), face = "bold"),
  axis.text.y = element_text(size = rel(0.7), face = "bold"),
  axis.title.x = element_blank(),
  axis.title.y = element_text(size = rel(0.6), face = "bold"),
  axis.line = element_line(colour = "black", linewidth = 1.2),
  legend.key.width = unit(rel(0.5), "cm"),
  legend.key.height = unit(rel(0.5), "cm"),
  legend.text = element_text(size = rel(0.5), face = "bold")
  )

```

###HF prevalence

```{r}

#HF
hf_figure_labels <- c(
  "[A] Crude prevalence (proportion)",
  "[B] Crude prevalence (absolute number)",
  "[C] Age-standardized prevalence",
  "[D] Age-stratified prevalence"
)
ggarrange(trends_hf_fig + rescale_ggplot, 
          trends_hf_n_fig  + rescale_ggplot,
          trends_hf_aa_fig + rescale_ggplot, 
          hf_age_strat_fig + rescale_ggplot,
          labels = hf_figure_labels,
          ncol = 2, nrow = 2,
          font.label = list(size = 15),
          hjust = -0.01)

grid_2by2(width = 3)
hf_panels_fig <- grid.grab()
ggsave(hf_panels_fig,
       filename = paste0("Figures/overall/Figure 1.pdf"),
       width = 16,
       height = 9,
       dpi = 600)


```

###HTN

```{r}

#HTN
htn_figure_labels <- c(
  "[A] Elevated blood pressure (Stage 1 or 2 hypertension)",
  "[B] Stage 1 hypertension",
  "[C] Stage 2 hypertension",
  "[D] Cumulative prevalence"
)
ggarrange(trends_fig_rfs$overall$`elevated blood pressure` + rescale_ggplot, 
          trends_fig_rfs$overall$`stage 1 hypertension` + rescale_ggplot, 
          trends_fig_rfs$overall$`stage 2 hypertension` + rescale_ggplot,
          cum_figs$`elevated blood pressure`  + rescale_ggplot,
          labels = htn_figure_labels,
          ncol = 2, nrow = 2,
          font.label = list(size = 15),
          hjust = -0.01)

grid_2by2(width = 3)
htn_panels_fig <- grid.grab()
ggsave(htn_panels_fig,
       filename = paste0("Figures/overall/Figure 2.pdf"),
       width = 16,
       height = 9,
       dpi = 600)


```

###Obesity

```{r}
#Obesity
obesity_figure_labels <- c(
  "[A] Obesity (class I-III)",
  "[B] Class I/II obesity",
  "[C] Class III obesity",
  "[D] Cumulative prevalence"
)
ggarrange(trends_fig_rfs$overall$obesity + rescale_ggplot, 
          trends_fig_rfs$overall$`class I or II obesity` + rescale_ggplot, 
          trends_fig_rfs$overall$`class III obesity` + rescale_ggplot,
          cum_figs$obesity  + rescale_ggplot,
          labels = obesity_figure_labels,
          ncol = 2, nrow = 2,
          font.label = list(size = 15),
          hjust = -0.01)

grid_2by2(width = 3)
obesity_panels_fig <- grid.grab()
ggsave(obesity_panels_fig,
       filename = paste0("Figures/overall/Figure 3.pdf"),
       width = 16,
       height = 9,
       dpi = 600)

```

###Diabetes

```{r}
#DM
obesity_figure_labels <- c(
  "[A] Diabetes or prediabetes",
  "[B] Prediabetes",
  "[C] Diabetes",
  "[D] Cumulative prevalence"
)
ggarrange(trends_fig_rfs$overall$`diabetes or pre-diabetes` + rescale_ggplot, 
          trends_fig_rfs$overall$prediabetes + rescale_ggplot, 
          trends_fig_rfs$overall$diabetes + rescale_ggplot,
          cum_figs$`insulin resistance`  + rescale_ggplot,
          labels = obesity_figure_labels,
          ncol = 2, nrow = 2,
          font.label = list(size = 15),
          hjust = -0.01)

grid_2by2(width = 3)
dm_panels_fig <- grid.grab()
ggsave(dm_panels_fig,
       filename = paste0("Figures/overall/Figure 4.pdf"),
       width = 16,
       height = 9,
       dpi = 600)

```

###Cholesterol/MI/CKD

```{r}
#Figure 5 labels
figure5_labels <- c(
  "[A] Hypercholesterolemia",
  "[A] Elevated LDL-C",
  "[C] Myocardial infarction",
  "[D] Chronic kidney disease"
)
ggarrange(trends_fig_rfs$overall$hypercholesterolemia + rescale_ggplot, 
          trends_fig_rfs$overall$`elevated low-density lipoprotein cholesterol` + rescale_ggplot, 
          trends_fig_rfs$overall$`a previous myocardial infarction` + rescale_ggplot, 
          trends_fig_rfs$overall$`chronic kidney disease` + rescale_ggplot,
          labels = figure5_labels,
          ncol = 2, nrow = 2,
          font.label = list(size = 15),
          hjust = -0.01)

grid_2by2(width = 3)
f5_panels_fig <- grid.grab()
ggsave(f5_panels_fig,
       filename = paste0("Figures/overall/Figure 5.pdf"),
       width = 16,
       height = 9,
       dpi = 600)

```

##Difference in ORs

###Model


```{r}
#Load some packages necessary for this part.
library(emmeans)
library(rms)

#Function to round P-values (of interaction)
p <- function(p) ifelse(p < 0.001, "<0.001", round(p, 3) %>% format(nsmall = 3))

#Function to get ORs for HF and non-HF and difference in ORs
or_diff <- function(outcome, survey_design_name) {
  
  
  #Set formula
  reg_formula <- as.formula(paste0(outcome, " ~ hf*year + rcs(age, 5) + sex"))
#Regress
svyglm(
  design = get(survey_design_name),
  family = binomial,
  formula = reg_formula) -> m

#Get HF-stratified ORs
emmeans(m, 
        data = full_data,
        trt.vs.ctrl ~ year|hf,#Contrasts of interest
        at = list(year = c(1980, 1990), #OR per 10-year change
                  hf = c("No", "Yes"))) -> comp

#Capture double differences
comp %>% 
  pairs %>%
  pairs(by = NULL) %>%
  confint %>%
  data.frame %>%
  rename(any_of(
                c("dd_pe" = "estimate",
                  "dd_lci" = "asymp.LCL",
                  "dd_uci" = "asymp.UCL"))) %>%
              mutate(across(.cols = c("dd_pe", "dd_lci", "dd_uci"), #Exponentiate
                  ~ . %>% exp)) %>%
  select(dd_pe, dd_lci, dd_uci) %>%
  data.table -> dd_df



#Collect in a data.frame
comp$contrasts %>%
  confint(adjust = "none") %>%
  data.frame %>%
  rename(any_of(
                c("pe" = "estimate",
                  "lci" = "asymp.LCL",
                  "uci" = "asymp.UCL"))) %>%
              mutate(across(.cols = c("pe", "lci", "uci"), #Exponentiate
                  ~ . %>% exp)) %>%
  select(hf, pe, lci, uci) %>%
  data.table -> comp


#Add p-value for interaction
comp[, pint := regTermTest(m, "hf:year")$p %>% as.numeric %>% p]

#Assign outcome
comp[, outcome := outcome]

#Create PE-CI
comp[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]


#Flip if necessary (for reductions)
if(all(comp$pe < 1)) {
  dd_df[, dd_pe := 1/dd_pe]
  dd_df[, (c("dd_uci", "dd_lci")) := .(1/dd_lci, 1/dd_uci)]
  }

#Combine with dd_df
comp <- cbind(comp, dd_df)

#Create PE-CI for double differences
comp[, dd_pe_ci := paste0(r2(dd_pe), " (", r2(dd_lci), " to ", r2(dd_uci), ")")]

#Return
comp
}


#Loop over desired outcomes
mapply(or_diff, outcome = c("htn", "obesity", "dm",
                            "mi", "ckd345_albumin", 
                            "hypercholesterol", "hyperldl", "hypertg"), 
       survey_design_name = c("sdata", "sdata", "sdata",
                              "sdata_q", "sdata", 
                              "sdata", "sdata_fast", "sdata_fast"),
       SIMPLIFY = F) -> or_diff_df
or_diff_df <- rbindlist(or_diff_df)


```

###Forest plot

#### Function to beautify forest

```{r}
#Get aesthetics for forest plot
forestploter_aes <- function(forest = forest, fp_df,
                             col_names = TRUE, specific_rows = TRUE,
                             marker = "header") {
  ##Header aesthetics
  if(col_names == TRUE) {
    
    #Underline headers
    forest <- add_underline(plot = forest, part = "header")
    
    # Bold column headers text
    forest <- edit_plot(forest,
                        part = "header",
                        gp = gpar(fontface = "bold"))
  }
  
  #Specific row aesthetics
  if(specific_rows == TRUE) {
    # Bold specific rows
    forest <- edit_plot(forest,
                        part = "body",
                        gp = gpar(fontface = "bold"),
                        row = fp_df[, marker] %>% pull %>% which)
    
    forest <- add_underline(forest, 
                            row = c(fp_df[, marker] %>% pull %>% which, 
                                    fp_df[, marker] %>% pull %>% which %>% {.-1}))
    
      }
  
}

```

#### Function to create forest

```{r}

#Grouped forest plot
grouped_fp <- function(forest_df, variable, grouping_variable,
                       title,
                       footnote,
                       space,
                       n_rows,
                       es_label,
                       xlog,
                       xlimits = c(0.4, 2.5), 
                       xticks = c(0.4, 0.5, 1, 2, 2.5),
                       size = 13) {
  
  #Set theme
  fp_theme <- forest_theme(base_size = size,
                           title_cex = 1.5,
                           footnote_cex = 0.7,
                           refline_lwd = 2,
                           xaxis_lwd = 2,
                           xaxis_cex = 1,
                           ci_pch = 15,
                           ci_Theight = 0.25,
                           ci_col = "black",
                           title_just = "center",
                           colhead = list(
                             fg_params = list(hjust = 0.5, 
                                              x = c(0.5, 0.5, 0.5)))
  )
  
  
  #Arrange according to outcomes
  forest_df <- arrange(forest_df, get(grouping_variable))
  
  #Identify unique grouping variable indices
  grouping_indices <- which(forest_df[, get(grouping_variable)] %>% {!duplicated(.)})
  
  #Lapply to create header rows
  lapply(grouping_indices,
         function(i)
           rbind(tibble(!!as.symbol(grouping_variable) := forest_df[i, get(grouping_variable)]) %>% data.table, 
                 forest_df[i:(i + (n_rows - 1) ),],
                 fill = T)
  ) %>% rbindlist -> forest_df
  
  
  #Combine variable and grouping_variable in same column
  forest_df[, var := fifelse(is.na(pe_ci), get(grouping_variable) %>% as.character, get(variable))]
  
  forest_df[!is.na(pint), pint %>% unique, by = get(grouping_variable)][, V1] -> pints
  
  #Assign P for interaction
  forest_df[, pint := NULL]
  forest_df[is.na(pe_ci), pint := pints]
  forest_df[!is.na(pe_ci), pint := ""]
  
  #Assign headings
  forest_df[, heading := fifelse(is.na(pe_ci), TRUE, FALSE)]
  
  #Pad
  forest_df[, var := paste0(var, "      ")]
  forest_df[!is.na(pe_ci), var := paste0("      ", var)]
  
  
  #Remake into data.frame
  data.frame(forest_df) -> forest_df
  #Replace NAs
  forest_df[is.na(forest_df)] <- ""
  
  #Insert empty column for FP plotting space
  forest_df$"space" <- paste(rep(" ", space), collapse = " ")
  
  #Columns to get
  fp_cols <- c(
    which(colnames(forest_df) == "var"),
    length(forest_df),
    which(colnames(forest_df) == "pe_ci"),
    which(colnames(forest_df) == "pint"),
    which(colnames(forest_df) == "dd_pe_ci")
  )
  
  
  #Forest
  forest <- forest(data = forest_df %>%
                     dplyr::select(fp_cols) %>%
                     rename("Variable/outcome" = var,
                            "  " = space,
                            !!as.symbol(es_label) := pe_ci,
                            "P-value for interaction" = pint,
                            "Ratio of ratios (95% CI)**" = dd_pe_ci),
                   #Labels and est/ci
                   est = forest_df$pe  %>% as.numeric,
                   ci_column = 2,
                   lower = forest_df$lci %>% as.numeric,
                   upper = forest_df$uci %>% as.numeric,
                   #Box Aesthetics,
                   sizes = 5,
                   #X-axis
                   ticks_at = xticks,
                   xlog = xlog,
                   xlim = xlimits,
                   #Title,
                   title = title,
                   footnote = footnote,
                   #Theme
                   theme = fp_theme
  )
  
  library(grid)
  forest_df <- tibble(forest_df)
  forestploter_aes(forest, fp_df = forest_df, marker = "heading") -> forest
  forest
  
}


```

#### Plot

```{r}
#Load forestploter package
library(forestploter)

#Coerce to character
or_diff_df[, hf := as.character(hf)]

#Replace outcome names
c(
   "htn" = "Elevated blood pressure",
   "obesity" = "Obesity",
  "dm" = "Diabetes",
  "hypercholesterol" = "Hypercholesterolemia",
  "hyperldl" = "Elevated LDL Cholesterol",
  "hypertg" = "Hypertriglyceridemia",
  "^mi$" = "Myocardial Infarction",
  "ckd345_albumin" = "Chronic kidney disease"
) -> replacement_string
or_diff_df[, outcome := str_replace_all(outcome, replacement_string)]


#Reorder
or_diff_df[, outcome]
or_diff_df[, outcome := factor(outcome, levels = replacement_string)]

#Remove second ratio of ratios row
or_diff_df[seq(1, nrow(or_diff_df), 2), dd_pe_ci := ""]

#Replace yes/no
or_diff_df[, hf := str_replace_all(hf, c("Yes"  =  "HF present",
                                          "No" = "HF absent"))]

#Forest
grouped_fp(
  forest_df = or_diff_df,
  variable = "hf", grouping_variable = "outcome",
  title = "Differences in temporal trends of risk factors between respondents with and without heart failure",
  footnote = "\n*Odds ratios (OR) are per 10-year increments and are adjusted for age and sex.\n**This ratio quantifies the extent to which a given increase or decrease was greater in degree among respondents with heart failure compared to respondents without heart failure",
  n_rows = 2,
  space = 50,
  es_label = "Odds ratio (95% CI)*",
  size = 15,
  xlog = T
) -> forest_fig_binary

#Save
ggsave(filename = "Figures/Overall/Forest Figure [Binary].pdf",
       forest_fig_binary,
       dpi = 600,
       width = 16, height = 9
       )

```

##Mean differences

###Model

```{r}


# outcome <- "bmi"
# survey_design_name <- "sdata"
#Function to get ORs for HF and non-HF and difference in MDs
mean_diff <- function(outcome, survey_design_name) {
  
  
  #Set formula
  reg_formula <- as.formula(paste0(outcome, " ~ hf*year + rcs(age, 5) + sex"))
#Regress
svyglm(
  design = get(survey_design_name),
  formula = reg_formula) -> m

#Get HF-stratified MDs
emmeans(m, 
        data = full_data,
        trt.vs.ctrl ~ year|hf,#Contrasts of interest
        at = list(year = c(1980, 1990), #MD per 10-year change
                  hf = c("No", "Yes"))) -> comp

#Capture double differences
comp %>% 
  pairs %>%
  pairs(by = NULL) %>%
  confint %>%
  data.frame %>%
  rename(any_of(
                c("dd_pe" = "estimate",
                  "dd_lci" = "lower.CL",
                  "dd_uci" = "upper.CL"))) %>%
  select(dd_pe, dd_lci, dd_uci) %>%
  data.table -> dd_df

#Collect in a data.frame
comp$contrasts %>%
  confint(adjust = "none") %>%
  data.frame %>%
  rename(any_of(
                c("pe" = "estimate",
                  "lci" = "lower.CL",
                  "uci" = "upper.CL"))) %>%
  select(hf, pe, lci, uci) %>%
  data.table -> comp


#Add p-value for interaction
comp[, pint := regTermTest(m, "hf:year")$p %>% as.numeric %>% p]

#Assign outcome
comp[, outcome := outcome]

#Create PE-CI
comp[, pe_ci := paste0(r2(pe), " (", r2(lci), " to ", r2(uci), ")")]


#Flip if necessary (for reductions)
if(all(comp$pe < 0)) {
  dd_df[, dd_pe := -dd_pe]
  dd_df[, (c("dd_uci", "dd_lci")) := .(-dd_lci, -dd_uci)]
  }

#Combine with dd_df
comp <- cbind(comp, dd_df)

#Create PE-CI for double differences
comp[, dd_pe_ci := paste0(r2(dd_pe), " (", r2(dd_lci), " to ", r2(dd_uci), ")")]

#Return
comp
}


#Loop over desired outcomes
mapply(mean_diff, outcome = c("sbp", "dbp",
                              "bmi",
                              "hba1c",
                              "tc", "ldl", "tg",
                              "egfr", "uacr"), 
       survey_design_name = c("sdata", "sdata",
                              "sdata",
                              "sdata",
                              "sdata", "sdata_fast", "sdata_fast",
                              "sdata", "sdata"),
       SIMPLIFY = F) -> mean_diff_df
mean_diff_df <- rbindlist(mean_diff_df)

```

###Tabulate

```{r}

#Coerce HF to character
mean_diff_df[, hf := as.character(hf)]

#Replace outcome names
c(
   "sbp" = "Systolic blood pressure (mmHg)",
   "dbp" = "Diastolic blood pressure (mmHg)",
   "bmi" = "Body mass index (kg/m)",
  "hba1c" = "HbA1c (%)",
  "tc" = "Total cholesterol (mg/dL)",
  "ldl" = "LDL Cholesterol (mg/dL)",
  "tg" = "Triglycerides (mg/dL)",
  "egfr" = "eGFR (mL/min/1.73m)",
  "uacr" = "Urine albumin-creatinine ratio (mg/g)"
) -> replacement_string
mean_diff_df[, outcome := str_replace_all(outcome, replacement_string)]


#Reorder
mean_diff_df[, outcome := factor(outcome, levels = replacement_string)]

#Remove second ratio of ratios row
mean_diff_df[seq(1, nrow(mean_diff_df), 2), dd_pe_ci := ""]
mean_diff_df[seq(1, nrow(mean_diff_df), 2), pint := ""]

#Replace yes/no
mean_diff_df[, hf := str_replace_all(hf, c("Yes"  =  "HF present",
                                          "No" = "HF absent"))]

#Forest
# grouped_fp(
#   forest_df = mean_diff_df,
#   variable = "hf", grouping_variable = "outcome",
#   title = "Average temporal difference in average levels of different risk markers in respondents with or without heart failure",
#   footnote = "\n*Mean differences are per 10-year increments and are adjusted for age and sex.\n**This ratio quantifies the extent to which a given increase or decrease was greater in degree among respondents with heart failure compared to respondents without heart failure",
#   n_rows = 2,
#   space = 50,
#   size = 15,
#   xlog = F,
#   es_label = "Mean difference (95% CI)*",
# ) -> forest_fig_cont

#Table
mean_diff_df %>%
  select(outcome, hf, pe_ci, pint, dd_pe_ci) %>%
  rename(Outcome = outcome,
         `Heart failure` = hf,
         `Mean difference (95% CI)` = pe_ci,
         `P-value for interaction` = pint,
         `Excess difference in respondents with HF (95% CI)` = dd_pe_ci) %>%
  flextable %>%
  merge_v(j = 1) %>%
  bold(j = 1) %>%
  bold(part = "header") %>%
  border_inner %>% border_outer %>%
  set_caption("eTable 5. Temporal difference in average levels of different risk factors between respondents with and without heart failure") %>%
  footnote(i = 1, j = c(3, 5), part = "header",
           ref_symbols = "*",
           value = as_paragraph("Differences are per 10-year increments and are adjusted for sex and age. Excess difference represents how much more or less pronounced a given increase or decrease was in respondents with heart failure vs respondents without heart failure.")) %>%
  width(width = 1.5) %>%
  save_as_docx(path = "Tables/eTable 5.docx")

```

## Create density figures

### Function

```{r}
#Function to plot density
plot_density <- function(outcome_var, outcome_label,
                                 weight, 
                                 survey_cycles1, survey_cycles2) {


  #Filter to necessary columns
  full_data[hf == "Yes" & !is.na(get(outcome_var)) &
              survey_cycle %in% c(survey_cycles1, survey_cycles2),
            .(outcome = get(outcome_var), survey_cycle, w = get(weight))
            ] -> density_data
  
  #Rename survey cycles
  #First 6 years
  density_data[survey_cycle %in% survey_cycles1,
               cycle := paste0(
                 min(unlist(str_split(survey_cycles1, pattern = " to "))),
                 " to ",
                 max(unlist(str_split(survey_cycles1, pattern = " to ")))
               )]
  
  #Last 6 years
  density_data[survey_cycle %in% survey_cycles2,
               cycle := paste0(
                 min(unlist(str_split(survey_cycles2, pattern = " to "))),
                 " to ",
                 max(unlist(str_split(survey_cycles2, pattern = " to ")))
               )]
  
  #We will retrieve the maximum y-axis value to adjust the positioning of the box plot below
  ggplot_build(
    ggplot(data = density_data,
           aes(x = outcome, y = after_stat(density), weight = w, fill = cycle)) +
      geom_density()
  )$data[[1]]$density %>%
    max -> max_density
  
  
  #Plot
  ggplot(data = density_data,
       aes(x = outcome,
           y = after_stat(density),
           fill = cycle,
           weight = w)
       ) +
  geom_density(alpha = 0.5, lwd = 1.1) +
    geom_boxplot(inherit.aes = F,
                 position = position_dodge(width = max_density/10),
                 aes(x = outcome, y = -max_density/15, fill = cycle, weight = w),
                 alpha = 0.5, lwd = 1.1, width = max_density/12.5,
                 outlier.color = NA,
                 show.legend = F) +
    #Scales
  scale_color_jama(name = "Survey cycle") +
  scale_fill_jama(name = "Survey cycle") +
  scale_x_continuous(name = outcome_label,
                     trans = ifelse(outcome_var == "uacr", "log", "identity"),
                     breaks = {if(outcome_var == "uacr") c(1, 10, 30, 100, 300, 1000, 3000) else breaks_pretty(n = 20)}
                     ) +
    scale_y_continuous(name = "Density",
                       expand = c(0.01, 0)) +
  theme_bw() +
  theme(text = element_text(size = 23, face = "bold"),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", size = 0),
        legend.position = c(0.8, 0.8),
        legend.key.width = unit(4, "cm"),
        legend.key.height = unit(0.8, "cm")
  ) -> density_fig

#Save
ggsave(density_fig,
       filename = paste0("Figures/overall/Density figure for ", outcome_var, ".pdf"),
       dpi = 600,
       width = 16, height = 9)

#Return
plot(density_fig)

}

```

### Loop over outcomes

```{r}

#Create a list containing variables needed to loop the density function
list(outcome_vars = c("sbp", "dbp",
                      "bmi",
                      "hba1c",
                      "tc", "ldl", "tg",
                      "egfr", "uacr"),
     outcome_labels = c("Systolic blood pressure (mmHg)",
                       "Diastolic blood pressure (mmHg)",
                       "Body mass index (kg/m)",
                      "HbA1c (%)",
                      "Total cholesterol (mg/dL)",
                      "LDL Cholesterol (mg/dL)",
                       "Triglycerides (mg/dL)",
                      "eGFR (mL/min/1.73m)",
                      "Urine albumin-creatinine ratio (mg/g)"),
     weights = c("exam_weight", "exam_weight",
                "exam_weight",
                "exam_weight",
                "exam_weight", "fast_weight", "fast_weight",
                "exam_weight", "exam_weight"),
     survey_cycles1 = list(c("1988 to 1991", "1991 to 1994")) %>% rep(9),
     survey_cycles2 = c(list(c("2017 to 2020", "2021 to 2023")) %>% rep(5),
                        list(c("2015 to 2016", "2017 to 2020")) %>% rep(4)
                        )
     ) -> density_list



#Loop over desired outcomes
mapply(plot_density,
       outcome_var = density_list$outcome_vars, 
       outcome_label = density_list$outcome_labels,
       weight = density_list$weights,
       survey_cycles1 = density_list$survey_cycles1,
       survey_cycles2 = density_list$survey_cycles2,
       SIMPLIFY = FALSE
       ) -> density_figs
names(density_figs) <- density_list$outcome_vars

```

##Trends within disease states

###Define a theme to add on top

```{r}
theme_bw() +
  theme(text = element_text(size = 23),
        plot.title=element_text(face = "bold",hjust = 0.0, size = 18),
        plot.subtitle = element_text(face = "bold", size = 10, hjust = 0.0, color = "grey45"),
        axis.text.x = element_text(size = 20, face = "bold"),
        axis.text.y = element_text(size = 20, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.line = element_line(colour = "black", linewidth = 1.2),
        plot.margin = margin(0.5, 1, 0.5, 1, "cm"),
        strip.text = element_text(size = 13, face = "bold"),
        legend.position = "top") -> cont_theme
```

###BMI

####Model

```{r}
svyglm(
  design = subset(sdata, bmi < 30 & hf == "Yes"),
  bmi ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_bmi_cont1
svyglm(
  design = subset(sdata, bmi >= 30 & hf == "Yes"),
  bmi ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_bmi_cont2

lapply(list(m_bmi_cont1, m_bmi_cont2),
       svyglm_preds
       ) -> m_bmi_preds
m_bmi_preds[[1]][, label := "No Obesity"]
m_bmi_preds[[2]][, label := "Obesity"] 
m_bmi_preds <- rbindlist(m_bmi_preds)


```

####Plot

```{r}
ggplot(data = m_bmi_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "BMI (kg/m)",
    breaks = seq(0, 100, 1)
  ) +
  #Add title
  ggtitle(paste0("Mean BMI in patients with heart failure, according to whether BMI meets (or not) the threshold for obesity, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> bmi_cont_fig

ggsave(bmi_cont_fig,
       filename = "Figures/overall/Continuous BMI.pdf",
       dpi = 600, width = 16, height = 9
       )

```

###HbA1c

####Model
```{r}
svyglm(
  design = subset(sdata, hba1c < 6.5 & hf == "Yes"),
  hba1c ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_hba1c_cont1
svyglm(
  design = subset(sdata, hba1c >= 6.5 & hf == "Yes"),
  hba1c ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_hba1c_cont2

lapply(list(m_hba1c_cont1, m_hba1c_cont2),
       svyglm_preds
       ) -> m_hba1c_preds
m_hba1c_preds[[1]][, label := "No diabetes"] 
m_hba1c_preds[[2]][, label := "Diabetes"] 
m_hba1c_preds <- rbindlist(m_hba1c_preds)

```

####Plot

```{r}
ggplot(data = m_hba1c_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "HbA1c (%)"
  ) +
  #Add title
  ggtitle(paste0("Mean HbA1c levels in patients with heart failure, according to whether HbA1c meets (or not) the threshold for DM, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> hba1c_cont_fig

ggsave(hba1c_cont_fig,
       filename = "Figures/overall/Continuous HbA1c.pdf",
       dpi = 600, width = 16, height = 9
       )

```

###HTN

#####SBP

######Model

```{r}
svyglm(
  design = subset(sdata, (dbp < 80 & sbp < 130) & hf == "Yes"),
  sbp ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_sbp_cont1
svyglm(
  design = subset(sdata, (dbp >= 80 | sbp >= 130) & hf == "Yes"),
  sbp ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_sbp_cont2

lapply(list(m_sbp_cont1, m_sbp_cont2),
       svyglm_preds
       ) -> m_sbp_preds
m_sbp_preds[[1]][, label := "No hypertension"]
m_sbp_preds[[2]][, label := "Hypertension"] 
m_sbp_preds <- rbindlist(m_sbp_preds)

```

######Plot

```{r}
ggplot(data = m_sbp_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Systolic blood pressure (mmHg)",
    breaks = seq(0, 300, 2)
  ) +
  #Add title
  ggtitle(paste0("Mean systolic blood pressure levels in patients with heart failure, in patients with or without hypertension, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> sbp_cont_fig

ggsave(sbp_cont_fig,
       filename = "Figures/overall/Continuous SBP.pdf",
       dpi = 600, width = 16, height = 9
       )

```

#### DBP

###### Model

```{r}
svyglm(
  design = subset(sdata, (dbp < 80 & dbp < 130) & hf == "Yes"),
  dbp ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_dbp_cont1
svyglm(
  design = subset(sdata, (dbp >= 80 | dbp >= 130) & hf == "Yes"),
  dbp ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_dbp_cont2

lapply(list(m_dbp_cont1, m_dbp_cont2),
       svyglm_preds
) -> m_dbp_preds
m_dbp_preds[[1]][, label := "No hypertension"]
m_dbp_preds[[2]][, label := "Hypertension"] 
m_dbp_preds <- rbindlist(m_dbp_preds)

```

###### Plot

```{r}
ggplot(data = m_dbp_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Diastolic blood pressure (mmHg)",
    breaks = seq(0, 300, 2)
  ) +
  #Add title
  ggtitle(paste0("Mean diastolic blood pressure levels in patients with heart failure, in patients with or without hypertension, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> dbp_cont_fig

ggsave(dbp_cont_fig,
       filename = "Figures/overall/Continuous DBP.pdf",
       dpi = 600, width = 16, height = 9
       )
```


###Hypercholesterol

####Model

```{r}
svyglm(
  design = subset(sdata, tc < 200 & hf == "Yes"),
  tc ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_tc_cont1
svyglm(
  design = subset(sdata, tc >= 200 & hf == "Yes"),
  tc ~ rcs(year, parms = c(1990.25, 2004.00, 2012.00, 2022.70))) -> m_tc_cont2

lapply(list(m_tc_cont1, m_tc_cont2),
       svyglm_preds
) -> m_tc_preds
m_tc_preds[[1]][, label := "No Hypercholesterolemia"] 
m_tc_preds[[2]][, label := "Hypercholesterolemia"] 
m_tc_preds <- rbindlist(m_tc_preds)

```
####Plot
```{r}

ggplot(data = m_tc_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "Total cholesterol (mg/dL)",
    breaks = seq(0, 300, 10)
  ) +
  #Add title
  ggtitle(paste0("Mean total cholesterol levels in patients with heart failure, in patients with or without hypercholesterolemia, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> tc_cont_fig

ggsave(tc_cont_fig,
       filename = "Figures/overall/Continuous TC.pdf",
       dpi = 600, width = 16, height = 9
       )
```

###LDL

####Model

```{r}
svyglm(
  design = subset(sdata_fast, ldl < 160 & hf == "Yes"),
  ldl ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_ldl_cont1
svyglm(
  design = subset(sdata_fast, ldl >= 160 & hf == "Yes"),
  ldl ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_ldl_cont2

lapply(list(m_ldl_cont1, m_ldl_cont2),
       svyglm_preds, newdata = data.frame(year = seq(1988, 2020, 1))
) -> m_ldl_preds
m_ldl_preds[[1]][, label := "LDL-C less than 160mg/dL"] 
m_ldl_preds[[2]][, label := "LDL-C 160mg/dL or greater"] 
m_ldl_preds <- rbindlist(m_ldl_preds)

```

####Plot

```{r}
ggplot(data = m_ldl_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
       ) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "LDL-C (mg/dL)",
    breaks = seq(0, 300, 10)
  ) +
  #Add title
  ggtitle(paste0("Mean LDL-C levels in patients with heart failure, in patients with or without LDL-C greater than 160mg/dL, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> ldl_cont_fig

ggsave(ldl_cont_fig,
       filename = "Figures/overall/Continuous LDL-C.pdf",
       dpi = 600, width = 16, height = 9
       )

```

###CKD

####eGFR

#####Model

```{r}
svyglm(
  design = subset(sdata, egfr < 60 & hf == "Yes"),
  egfr ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_egfr_cont1
svyglm(
  design = subset(sdata, egfr >= 60 & hf == "Yes"),
  egfr ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_egfr_cont2

lapply(list(m_egfr_cont1, m_egfr_cont2),
       svyglm_preds, newdata = data.frame(year = seq(1988, 2020, 1))
) -> m_egfr_preds
m_egfr_preds[[1]][, label := "eGFR less than 60mL/min/1.73m)"] 
m_egfr_preds[[2]][, label := "eGFR 60mL/min/1.73m or greater)"] 
m_egfr_preds <- rbindlist(m_egfr_preds)

```

#####Plot

```{r}
ggplot(data = m_egfr_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "eGFR (mL/min/1.73m)",
    breaks = seq(0, 1000, 2)
  ) +
  #Add title
  ggtitle(paste0("Mean eGFR levels in patients with heart failure, in patients with or without eGFR <60mL/min/1.73m, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> egfr_cont_fig

ggsave(egfr_cont_fig,
       filename = "Figures/overall/Continuous eGFR.pdf",
       dpi = 600, width = 16, height = 9
)

```

####UACR

#####Model

```{r}
svyglm(
  design = subset(sdata, uacr < 30 & hf == "Yes"),
  uacr ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_uacr_cont1
svyglm(
  design = subset(sdata, uacr >= 30 & hf == "Yes"),
  uacr ~ rcs(year, parms = c(1990.25, 2008.00, 2018.60))) -> m_uacr_cont2

lapply(list(m_uacr_cont1, m_uacr_cont2),
       svyglm_preds, newdata = data.frame(year = seq(1988, 2020, 1))
) -> m_uacr_preds
m_uacr_preds[[1]][, label := "No Albuminuria"] 
m_uacr_preds[[2]][, label := "Albuminuria"] 
m_uacr_preds <- rbindlist(m_uacr_preds)
```

#####Plot

```{r}
ggplot(data = m_uacr_preds,
       aes(
         x = year,
         y = pe,
         ymax = uci,
         ymin = lci,
         color = label,
         fill = label
       )
) +
  #Points and lines and ribbons
  geom_point(size = 4) +
  geom_line(lwd = 1.5) +
  geom_ribbon(alpha = 0.2) +
  facet_grid(rows = vars(label), scales = "free_y") +
  #Colors
  scale_fill_jama(name = NULL) +
  scale_color_jama(name = NULL) +
  #Axes
  scale_x_continuous(
    name = "Year",
    breaks = seq(0, 5000, 2),
    expand = c(0.02, 0.0)
  ) +
  scale_y_continuous(
    name = "eGFR (mL/min/1.73m)"
  ) +
  #Add title
  ggtitle(paste0("Mean UACR levels in patients with heart failure, in patients with or without microalbuminuria, 1988-2023"),
          subtitle = "Line represents means across time and ribbons represent the 95% CI.") +
  #Theme
  cont_theme -> uacr_cont_fig

ggsave(uacr_cont_fig,
       filename = "Figures/overall/Continuous UACR.pdf",
       dpi = 600, width = 16, height = 9
)
```

## Table 1 (Results summary)

### Create data

```{r}
data.table(
  `Risk factor` = c(
    "Elevated blood pressure",
    "     Stage 1 hypertension",
    "     Stage 2 hypertension",
    "Obesity",
    "     Class I/II obesity",
    "     Class III obesity",
    "Impaired glucose homeostasis",
    "     Prediabetes",
    "     Diabetes",
    "Hypercholesterolemia",
    "     Elevated LDL-C",
    "     Hypertriglyceridemia",
    "Myocardial Infarction",
    "Chronic Kidney Disease",
    "     Reduced eGFR",
    "     Albuminuria"
  ),
  `Estimated % prevalence in 1988 (95% CI)` = c(
    htn_trends$overall$htn$preds_df[year == 1988, pe_ci],
    htn_trends$overall$htn1$preds_df[year == 1988, pe_ci],
    htn_trends$overall$htn2$preds_df[year == 1988, pe_ci],
    obesity_trends$overall$obesity$preds_df[year == 1988, pe_ci],
    obesity_trends$overall$obesity12$preds_df[year == 1988, pe_ci],
    obesity_trends$overall$obesity3$preds_df[year == 1988, pe_ci],
    dm_trends$overall$pre_dm_or_dm$preds_df[year == 1988, pe_ci],
    dm_trends$overall$pre_dm$preds_df[year == 1988, pe_ci],
    dm_trends$overall$dm$preds_df[year == 1988, pe_ci],
    hypercholesterol_trends$overall$preds_df[year == 1988, pe_ci],
    hyperldl_trends$overall$preds_df[year == 1988, pe_ci],
    hypertg_trends$overall$preds_df[year == 1988, pe_ci],
    mi_trends$overall$preds_df[year == 1988, pe_ci],
    ckd345_albumin_trends$overall$preds_df[year == 1988, pe_ci],
    ckd345_trends$overall$preds_df[year == 1988, pe_ci],
    albuminuria_trends$overall$preds_df[year == 1988, pe_ci]
  ),
  `Estimated % prevalence in 2023 or 2020 (95% CI)` = c(
    htn_trends$overall$htn$preds_df[year == 2023, pe_ci],
    htn_trends$overall$htn1$preds_df[year == 2023, pe_ci],
    htn_trends$overall$htn2$preds_df[year == 2023, pe_ci],
    obesity_trends$overall$obesity$preds_df[year == 2023, pe_ci],
    obesity_trends$overall$obesity12$preds_df[year == 2023, pe_ci],
    obesity_trends$overall$obesity3$preds_df[year == 2023, pe_ci],
    dm_trends$overall$pre_dm_or_dm$preds_df[year == 2023, pe_ci],
    dm_trends$overall$pre_dm$preds_df[year == 2023, pe_ci],
    dm_trends$overall$dm$preds_df[year == 2023, pe_ci],
    hypercholesterol_trends$overall$preds_df[year == 2023, pe_ci],
    hyperldl_trends$overall$preds_df[year == 2020, pe_ci],
    hypertg_trends$overall$preds_df[year == 2020, pe_ci],
    mi_trends$overall$preds_df[year == 2023, pe_ci],
    ckd345_albumin_trends$overall$preds_df[year == 2020, pe_ci],
    ckd345_trends$overall$preds_df[year == 2020, pe_ci],
    albuminuria_trends$overall$preds_df[year == 2020, pe_ci]
  ),
  `Prevalence ratio (95% CI)` = c(
    htn_trends$overall$htn$contrasts_df[, pe_ci],
    htn_trends$overall$htn1$contrasts_df[, pe_ci],
    htn_trends$overall$htn2$contrasts_df[, pe_ci],
    obesity_trends$overall$obesity$contrasts_df[, pe_ci],
    obesity_trends$overall$obesity12$contrasts_df[, pe_ci],
    obesity_trends$overall$obesity3$contrasts_df[, pe_ci],
    dm_trends$overall$pre_dm_or_dm$contrasts_df[, pe_ci],
    dm_trends$overall$pre_dm$contrasts_df[, pe_ci],
    dm_trends$overall$dm$contrasts_df[, pe_ci],
    hypercholesterol_trends$overall$contrasts_df[, pe_ci],
    hyperldl_trends$overall$contrasts_df[, pe_ci],
    hypertg_trends$overall$contrasts_df[, pe_ci],
    mi_trends$overall$contrasts_df[, pe_ci],
    ckd345_albumin_trends$overall$contrasts_df[, pe_ci],
    ckd345_trends$overall$contrasts_df[, pe_ci],
    albuminuria_trends$overall$contrasts_df[, pe_ci]
  ),
  `Prevalence ratio (95% CI) for women` = c(
    htn_trends$sex$htn$contrasts_df[int_factor == "Women", pe_ci],
    htn_trends$sex$htn1$contrasts_df[int_factor == "Women", pe_ci],
    htn_trends$sex$htn2$contrasts_df[int_factor == "Women", pe_ci],
    obesity_trends$sex$obesity$contrasts_df[int_factor == "Women", pe_ci],
    obesity_trends$sex$obesity12$contrasts_df[int_factor == "Women", pe_ci],
    obesity_trends$sex$obesity3$contrasts_df[int_factor == "Women", pe_ci],
    dm_trends$sex$pre_dm_or_dm$contrasts_df[int_factor == "Women", pe_ci],
    dm_trends$sex$pre_dm$contrasts_df[int_factor == "Women", pe_ci],
    dm_trends$sex$dm$contrasts_df[int_factor == "Women", pe_ci],
    hypercholesterol_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    hyperldl_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    hypertg_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    mi_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    ckd345_albumin_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    ckd345_trends$sex$contrasts_df[int_factor == "Women", pe_ci],
    albuminuria_trends$sex$contrasts_df[int_factor == "Women", pe_ci]
  ),
  `Prevalence ratio (95% CI) for men` = c(
    htn_trends$sex$htn$contrasts_df[int_factor == "Men", pe_ci],
    htn_trends$sex$htn1$contrasts_df[int_factor == "Men", pe_ci],
    htn_trends$sex$htn2$contrasts_df[int_factor == "Men", pe_ci],
    obesity_trends$sex$obesity$contrasts_df[int_factor == "Men", pe_ci],
    obesity_trends$sex$obesity12$contrasts_df[int_factor == "Men", pe_ci],
    obesity_trends$sex$obesity3$contrasts_df[int_factor == "Men", pe_ci],
    dm_trends$sex$pre_dm_or_dm$contrasts_df[int_factor == "Men", pe_ci],
    dm_trends$sex$pre_dm$contrasts_df[int_factor == "Men", pe_ci],
    dm_trends$sex$dm$contrasts_df[int_factor == "Men", pe_ci],
    hypercholesterol_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    hyperldl_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    hypertg_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    mi_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    ckd345_albumin_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    ckd345_trends$sex$contrasts_df[int_factor == "Men", pe_ci],
    albuminuria_trends$sex$contrasts_df[int_factor == "Men", pe_ci]
  ),
  `P-value for interaction by sex` = c(
    htn_trends$sex$p_interaction %>% p,
    "",
    "",
    obesity_trends$sex$p_interaction %>% p,
    "",
    "",
    dm_trends$sex$p_interaction %>% p,
    "",
    "",
    hypercholesterol_trends$sex$rrr_df$pval_int %>% p,
    "",
    "",
    mi_trends$sex$rrr_df$pval_int %>% p,
    ckd345_albumin_trends$sex$rrr_df$pval_int %>% p,
    "",
    ""
  ),
  `Prevalence ratio (95% CI) for Hispanic, Black, or other minority persons` = c(
    htn_trends$race_binary$htn$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    htn_trends$race_binary$htn1$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    htn_trends$race_binary$htn2$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    obesity_trends$race_binary$obesity$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    obesity_trends$race_binary$obesity12$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    obesity_trends$race_binary$obesity3$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    dm_trends$race_binary$pre_dm_or_dm$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    dm_trends$race_binary$pre_dm$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    dm_trends$race_binary$dm$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    hypercholesterol_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    hyperldl_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    hypertg_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    mi_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    ckd345_albumin_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    ckd345_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci],
    albuminuria_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic Black, Hispanic, or Other Minority", pe_ci]
  ),
  `Prevalence ratio (95% CI) for NH White persons` = c(
    htn_trends$race_binary$htn$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    htn_trends$race_binary$htn1$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    htn_trends$race_binary$htn2$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    obesity_trends$race_binary$obesity$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    obesity_trends$race_binary$obesity12$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    obesity_trends$race_binary$obesity3$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    dm_trends$race_binary$pre_dm_or_dm$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    dm_trends$race_binary$pre_dm$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    dm_trends$race_binary$dm$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    hypercholesterol_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    hyperldl_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    hypertg_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    mi_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    ckd345_albumin_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    ckd345_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci],
    albuminuria_trends$race_binary$contrasts_df[int_factor == "Non-Hispanic White", pe_ci]
  ),
  `P-value for interaction by race/ethnicity` = c(
    htn_trends$race_binary$p_interaction %>% p,
    "",
    "",
    obesity_trends$race_binary$p_interaction %>% p,
    "",
    "",
    dm_trends$race_binary$p_interaction %>% p,
    "",
    "",
    hypercholesterol_trends$race_binary$rrr_df$pval_int %>% p,
    "",
    "",
    mi_trends$race_binary$rrr_df$pval_int %>% p,
    ckd345_albumin_trends$race_binary$rrr_df$pval_int %>% p,
    "",
    ""
  )
  
) -> results_summary_table


```


### Tabulate

```{r}
results_summary_table %>%
  flextable %>%
  merge_h %>%
  bold(i = 1, part = "header") %>%
  bold(j = 1) %>%
  set_caption(caption = as_paragraph("Table 1. Summary of changes in the prevalence of risk factors among patients with heart failure in the US, 1988-2023.")) %>%
  flextable::width(width = 2) %>%
    footnote(i = 1, j = 3, part = "header",
           ref_symbols = "\U002A",
           value = as_paragraph("Data on low-density lipoprotein cholesterol (LDL-C), triglycerides, estimated glomerular filtration rate, and urine albumin-creatinine were only collected up to 2020 by NHANES.")) %>%
  footnote(i = 1, j = c(7, 10), part = "header",
           ref_symbols = "\U2020",
           value = as_paragraph("For this analysis, a P-value of 0.05 denoted statistical significance for all results, including interactions. Some practitioners use a more liberal threshold of 0.1 owing to the lesser statistical power of subgroup analyses.")) %>%
  footnote(i = 1, j = c(7, 10), part = "header",
           ref_symbols = "\U2021",
           value = as_paragraph("Because smooth interactions between time and race/sex were used, interaction P-values do not correspond to tests of differences in the reported prevalence ratios between subgroups but instead test for differences in in the shape of the trend across the entire study period.")) %>%
    footnote(i = 1, j = c(7, 10), part = "header",
           ref_symbols = "\U00A7",
           value = as_paragraph("To minimize the risk of false positive interactions with subgroup testing, interaction P-values for only the 6 main risk factor domains of interest (elevated blood pressure, impaired glucose homeostasis, obesity, hypercholesterolemia, myocardial infarction, and chronic kidney disease) are reported.")) %>%
  flextable::font(fontname = "Times New Roman") %>%
  border_inner %>%
  border_outer %>%
  save_as_docx(path = "Tables/Table 1.docx")
c("\U002A", "\U2020", "\U2021", "\U00A7")
```


## eTable 1 (N of respondents)

```{r}
data.table(
  `Survey Cycle` = full_data[, unique(survey_cycle)],
  `N with heart failure` = full_data[!is.na(hf), sum(hf == "Yes"), by = .(survey_cycle)][,V1],
  `N without heart failure` = full_data[!is.na(hf), sum(hf == "No"), by = .(survey_cycle)][,V1]
) -> etable1

flextable(etable1) %>%
  border_inner %>%
  border_outer %>%
  merge_h %>%
  bold(i = 1, part = "header") %>%
  bold(j = 1) %>%
  set_caption(caption = as_paragraph(paste0("eTable 1. Number of respondents, stratified by the presence or absence of heart failure, in each NHANES survey cycle."))) %>%
  flextable::font(fontname = "Times New Roman") %>%
  width(width = 2.5) %>%
  save_as_docx(path = "Tables/eTable 1.docx")



```


## eTable 2/3 (Baseline CCs)

###CCs function

```{r}

#Create a function to get a weighted median/IQR for continuous variables
wtd_median_iqr <- function(var, wt, rounding = 0) {
  paste0(round(wtd.quantile(var, weights = wt, na.rm = T)[3], rounding) %>% format(nsmall = rounding),
         " (", round(wtd.quantile(var, weights = wt, na.rm = T)[2], rounding) %>% format(nsmall = rounding),
         " to ",
         round(wtd.quantile(var, weights = wt, na.rm = T)[4], rounding) %>% format(nsmall = rounding),
         ")")
}

#Function to tabulate characteristics by survey cycle
table_survey_cycle_fn <- function(hf_subset, table_title, table_width, font_size) {
list(c("Age (years)", full_data[hf == hf_subset, wtd_median_iqr(age, int_weight), by = survey_cycle][, V1]),
     c("Proportion of women (%)", full_data[hf == hf_subset, sum(int_weight[sex == "Women"]/sum(int_weight)), by = survey_cycle][, r1(100*V1)]),
     c("Proportion of Non-Hispanic Black, Hispanic, or Other Minority (%)", full_data[hf == hf_subset, sum(int_weight[race_binary == "Non-Hispanic Black, Hispanic, or Other Minority"]/sum(int_weight)), by = survey_cycle][, r1(100*V1)]),
     c("Body mass index (kg/m)", full_data[hf == hf_subset & !is.na(bmi), wtd_median_iqr(bmi, exam_weight, 1), by = survey_cycle][, V1]),
     c("Systolic Blood pressure (mmHg)", full_data[hf == hf_subset & !is.na(sbp), wtd_median_iqr(sbp, exam_weight, 0), by = survey_cycle][, V1]),
     c("Diastolic Blood pressure (mmHg)", full_data[hf == hf_subset & !is.na(dbp), wtd_median_iqr(dbp, exam_weight, 0), by = survey_cycle][, V1]),
     c("HbA1c (%)", full_data[hf == hf_subset & !is.na(hba1c), wtd_median_iqr(hba1c, exam_weight, 1), by = survey_cycle][, V1]),
     c("Low-density lipoprotein cholesterol (mg/dL)", full_data[hf == hf_subset & !is.na(ldl), wtd_median_iqr(ldl, fast_weight, 0), by = survey_cycle][, V1]),
     c("Triglycerides (mg/dL)", full_data[hf == hf_subset & !is.na(ldl), wtd_median_iqr(tg, fast_weight, 0), by = survey_cycle][, V1]),
     c("Estimated glomerular filtration rate (mL/min/1.73m)", full_data[hf == hf_subset & !is.na(egfr), wtd_median_iqr(egfr, exam_weight, 0), by = survey_cycle][, V1]),
     c("Urine albumin-creatinine ratio (mg/g)", full_data[hf == hf_subset & !is.na(uacr), wtd_median_iqr(uacr, exam_weight, 0), by = survey_cycle][, V1]),
     c("Cardiovascular risk factors (%)", rep("", times = full_data[, uniqueN(survey_cycle)])),
     c("   Obesity", full_data[hf == hf_subset & !is.na(bmi), sum(exam_weight[bmi >= 30]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Class I/II Obesity", full_data[hf == hf_subset & !is.na(bmi), sum(exam_weight[bmi %between% c(30, 39.999)]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Class III Obesity", full_data[hf == hf_subset & !is.na(bmi), sum(exam_weight[bmi >= 40]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Prediabetes", full_data[hf == hf_subset & !is.na(hba1c), sum(exam_weight[hba1c >= 5.7 & hba1c < 6.5]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Prediabetes or diabetes", full_data[hf == hf_subset & !is.na(hba1c), sum(exam_weight[hba1c >= 5.7]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Prediabetes", full_data[hf == hf_subset & !is.na(hba1c), sum(exam_weight[hba1c %between% c(5.7, 6.4999)]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Diabetes", full_data[hf == hf_subset & !is.na(hba1c), sum(exam_weight[hba1c >= 6.5]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
      c("   Elevated Blood Pressure", full_data[hf == hf_subset & !is.na(sbp) & !is.na(dbp), sum(exam_weight[sbp >= 130 | dbp >= 80]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Stage 1 hypertension", full_data[hf == hf_subset & !is.na(sbp) & !is.na(dbp), sum(exam_weight[sbp %between% c(130, 139.999) | dbp %between% c(80, 89.9999)]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Stage 2 hypertension", full_data[hf == hf_subset & !is.na(sbp) & !is.na(dbp), sum(exam_weight[sbp >= 140 | dbp >= 90]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Hypercholesterolemia", full_data[hf == hf_subset & !is.na(tc) & !is.na(tg), sum(exam_weight[tc >= 200]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Hypertriglyceridemia", full_data[hf == hf_subset & !is.na(tg), sum(fast_weight[tg >= 200]/sum(fast_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   Elevated LDL-C", full_data[hf == hf_subset & !is.na(ldl), sum(fast_weight[ldl >= 200]/sum(fast_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   History of myocardial infarction (%)", full_data[hf == hf_subset & !is.na(mi), sum(int_weight[mi == "Yes"]/sum(int_weight)), by = survey_cycle][, r1(100*V1)]),
     c("   History of chronic kidney disease (%)", full_data[hf == hf_subset & !is.na(ckd345_albumin), sum(exam_weight[ckd345_albumin == "Yes"]/sum(exam_weight)), by = survey_cycle][, r1(100*V1)])
     ) -> table_list

#Add an N/A for variables which were not collected in the last survey cycle
lapply(table_list, function(x) if(length(x) < full_data[, uniqueN(survey_cycle) + 1]) c(x, "N/A") else x) -> table_list

#Unlist
unlist(table_list) -> table_vector
#Morph into a table
matrix(table_vector,
       ncol = full_data[, uniqueN(survey_cycle) + 1],
       byrow = T) %>%
  data.table -> table_dt

#Rename
names(table_dt) <- c("Variable", full_data[, unique(survey_cycle) %>% as.character])

#Flextable
flextable(table_dt) %>%
  border_inner %>%
  border_outer %>%
  merge_h %>%
  bold(i = 1, part = "header") %>%
  bold(j = 1) %>%
  set_caption(caption = as_paragraph(table_title)) %>%
  flextable::font(fontname = "Times New Roman") %>%
  fontsize(size = font_size) %>%
  width(., width = table_width) -> table_flex

#Return
table_flex
}

```

###Tabulate

```{r}
#Tabulate baseline CCs by survey cycle for respondents with HF
table_survey_cycle_fn(hf_subset = "Yes",
                      table_title = "eTable 2. Characteristics of respondents with heart failure in each NHANES survey cycle.",
                      table_width = 1.5,
                      font_size = 10) %>%
  save_as_docx(path = "Tables/eTable 2.docx")
#Then, those without HF
table_survey_cycle_fn(hf_subset = "No",
                      table_title = "eTable 3. Characteristics of respondents without heart failure in each NHANES survey cycle.",
                      table_width = 1.5,
                      font_size = 10) %>%
  save_as_docx(path = "Tables/eTable 3.docx")


```

## eTable 4 (HF Prevalence as a function of different variables)

### Functions to create plots

```{r}
library(gt)

t1_cont_ggplot <- function(pred_var, survey_weight) {
  #Declare formula
  reg_formula <- as.formula(paste0("hf", " ~ s(", pred_var, ", k = 4)"))
  
  #Run model
  gam(data = full_data,
      method = "REML",
      weight = full_data[, get(survey_weight)],
      family = "binomial",
      formula = reg_formula) -> reg_model
  
  #Predict
  pred_var_vector <- full_data[, get(pred_var) %>% unique %>% sort]
  predict(reg_model, 
          newdata = tibble(!!as.symbol(pred_var) := pred_var_vector)
  ) -> reg_preds
  reg_preds <- plogis(reg_preds)
  
  #Combine in a data.frame
  data.table(
    predictor = pred_var_vector,
    hf_prev = reg_preds
  ) -> gg_df
  
  #Trim to 5th and 95th percentiles
  trim_limits <- full_data[, quantile(get(pred_var), c(0.025, 0.975), na.rm = T)]
  gg_df <- gg_df[predictor %between% trim_limits]
  
  #Use log scale on X-axis for UACR
  trans_choice <- fifelse(pred_var == "uacr", "log", "identity")
  
  #Plot
  ggplot(data = gg_df,
         aes(y = hf_prev,
             x = predictor)
  ) +
    geom_line(lwd = 2.5) +
    scale_x_continuous(name = NULL,
                       breaks = {if(pred_var == "uacr") c(10, 100, 1000, 10000) else breaks_pretty(n = 5)},
                       trans = trans_choice) +
    scale_y_continuous(name = NULL,
                       labels = function(x) x * 100) +
    theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(),
          axis.text = element_text(size = 30, face = "bold")
    )
}

#For categorical variables
t1_cat_ggplot <- function(pred_var, survey_weight) {
  
  #Get proportions
  full_data[, .(n = sum(get(survey_weight))), 
            by = .(`predictor` = get(pred_var))][, .(predictor,
                                                     prop = n/sum(n))] -> gg_df
  
  #Plot
  ggplot(data = gg_df,
         aes(y = prop,
             x = predictor,
             fill = predictor)
  ) +
    geom_col(lwd = 2.5) +
    scale_x_discrete(name = NULL, labels = NULL) +
    scale_y_continuous(name = NULL,
                       labels = function(x) x * 100) +
    scale_fill_jama(name = NULL) +
    theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(),
          axis.text = element_text(size = 30, face = "bold"),
          legend.position = "top",
          legend.text = element_text(size = 40, face = "bold"),
          legend.key.width = unit(3, "cm")
    ) +
    guides(fill = guide_legend(nrow = 2, byrow = T))
  
}


#Function to create a mini ggplot2
t1_ggplot <- function(pred_var, survey_weight) {
  
  
  if(full_data[, is.numeric(get(pred_var))]) {
    t1_cont_ggplot(pred_var = pred_var, survey_weight = survey_weight)
  } else {
    t1_cat_ggplot(pred_var = pred_var, survey_weight = survey_weight)
  }
  

}

```

### Tabulate

```{r}
#Continuous variables in table 1
pred_vars <- c("age", 
               "sex",
               "race_binary",
                   "bmi", 
                   "sbp", "dbp",
                   "hba1c", "tc",
                   "ldl", "tg",
                   "egfr", "uacr")

#Their respective weights
survey_weights <- c("int_weight", 
                    "int_weight",
                    "int_weight",
                         "exam_weight",
                         "exam_weight", "exam_weight",
                         "exam_weight", "exam_weight", 
                         "fast_weight", "fast_weight",
                         "exam_weight", "exam_weight")

#loop over them to produce plots
mapply(t1_ggplot,
       pred_var = pred_vars,
       survey_weight = survey_weights,
       SIMPLIFY = F
       ) -> t1_plots
names(t1_plots) <- pred_vars
       
    
tibble(Variables = c("Age (years)", "Sex", "Race/ethnicity",
                     "Body mass index (kg/m)",
                "Systolic blood pressure (mmHg)", "Diastolic blood pressure (mmHg)",
                "HbA1c (%)", "Total cholesterol (mg/dL)",
                "Low-density lipoprotein cholesterol (mg/dL)", "Triglycerides (mg/dL)",
                "Estimated glomerular filtration rate (mL/min/1.73m)", "Urine albumin/creatinine ratio (mg/dL)"
                ),
       `Prevalence of heart failure (%)` = NA) -> new_t1

library(gt)
#Convert to gt and add plots
new_t1 %>%
  gt %>%
  text_transform(
    locations = cells_body(columns = `Prevalence of heart failure (%)`),
    fn = function(x) {
      t1_plots %>%
        ggplot_image(height = px(100),
                     aspect_ratio = 4)
    }
  ) -> new_t1


#Add title and bold
new_t1 %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_column_labels(everything()),
                     cells_body(),
                     cells_title()
    )
  ) -> new_t1

#Save
gtsave(new_t1, filename = "Tables/eTable 4.png")

```

# Save

```{r}
save(
  #Datasets
  full_data,
  sdata,
  #Trends
  hf_trends, aa_preds_df, year_rexvar, age_rexvar,
  htn_trends, htn_hist2_trends, bb_trends, acei_arb_trends, thiazide_trends, ccb_trends,
  obesity_trends, 
  dm_trends,
  hypercholesterol_trends, dyslipid_hist2_trends, hyperldl_trends, hypertg_trends, statin_trends,
  mi_trends,
  ckd345_trends, ckd345_albumin_trends, albuminuria_trends,
  #Multipanel figures
  hf_panels_fig, htn_panels_fig, obesity_panels_fig, dm_panels_fig,
  #HF Prevalence figure
  trends_hf_fig, trends_sex_hf_fig, trends_race_hf_fig, trends_hf_n_fig, hf_age_strat_fig, trends_hf_aa_fig,
   #RF Figures
  trends_fig_rfs,
  cum_figs,
  #Med figures
  trends_fig_meds,
  #Forest figure (binary)
  forest_fig_binary,
  #Density figures
  density_figs,
  #Trends within outcome categories
  bmi_cont_fig, hba1c_cont_fig, sbp_cont_fig, dbp_cont_fig,
  tc_cont_fig, ldl_cont_fig, egfr_cont_fig, uacr_cont_fig,
   #File
   file = "Rendering Data.RData")

```
